### 2022-05-11 README.Rmd
first commit

---
title: ohun_paper_analyses
author: AUTOR NAME
output: github_document
date: `r format(Sys.Date(), "%d-%m-%Y")`
editor_options:
  chunk_output_type: console
---

<!-- Short Description  -->
This is a short description of the project

*Updated on `r Sys.time()`*

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}

library(rmarkdown)

knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%"
)
```

## Table of contents
* [General info](#general-info)
  * [Analyses](#Analyses)
    * [To do list](#to-do-list)
      * [Contact](#contact)

## General info

General information about the project (study system, goals, etc)

## Analysis

* Analysis 1

## In progress

*  Something not finished yet

## To-do list

* Do the next analysis

## Status
Project is: _in progress_

## Contact
Created by [Autor name](website_URL)