### 2022-05-11 scripts/study_cases_bat_social_calls.Rmd

---
title: <center><font size="6"><b>Ohun</b></font></center>
subtitle: <center><font size="4"><b>Study cases</b></font></center>
author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas</a> & <a href="http:///">Gloriana Chaverri</a></font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    theme: default
    highlight: pygments
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

## Load packages
```{r packages, message=FALSE, warning = FALSE, echo = TRUE, eval = TRUE, include = TRUE}

## add 'developer/' to packages to be installed from github
x <- c("pbapply", "viridis", "knitr", "remotes", "maRce10/warbleR", "maRce10/ohun", "DT", "ranger")

aa <- lapply(x, function(y) {

  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]

  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  remotes::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  try(require(pkg, character.only = T), silent = T)
})

```

```{r set knitting options,  eval = TRUE, echo = FALSE}

opts_knit$set(dpi = 100, fig.width = 12, warning = FALSE, root.dir = "..", fig.align = 'center', message = FALSE)

```

## Set functions and global parameters
```{r functions, eval = TRUE, echo = TRUE}

cols <- viridis(10, alpha = 0.6)

warbleR_options(wav.path = "./data/raw")

```

# Bat calls data

## File descriptions

manual_detections_Raven_bats4testing.csv: Manual detection of all visible bat vocalizations (Raven)

manual_detections_bats4testing.csv: Manual detection of all high quality bat vocalizations (Avisoft)

automated_detections_bats4testing.csv: Automated detection with Eric's function

# Read data
```{r}

reference_all <- read.csv("./data/raw/manual_detections_Raven_bats4testing.csv")

reference_best <- read.csv("./data/raw/manual_detections_bats4testing.csv")

ericks_detection <- read.csv("./data/raw/automated_detections_bats4testing.csv")

```

# Check references

Exhaustive detection (all signals) 
```{r}
# all
check_sels(reference_all, pb = FALSE)
feature_reference(reference_all, path = "./data/raw")
```

High quality subset
```{r}

check_sels(reference_best, pb = FALSE)
feature_reference(reference_best, path = "./data/raw")
```

Detection from Erick's algorithm
```{r}

check_sels(ericks_detection, pb = FALSE)
feature_reference(ericks_detection, path = "./data/raw")

```

# Compare references (high quality vs exhaustive)

```{r}

# all vs best
diagnose_detection(reference = reference_all, detection = reference_best)

```
&nbsp;

<div class="alert alert-info">

- *High quality reference has ~70% of all signals*

</div>

&nbsp;

# Diagnose Erick's detection

Compare to exhaustive detection:

```{r}

diagnose_detection(reference = reference_all, detection = ericks_detection)

```
&nbsp;

<div class="alert alert-info">

- *Erick's algorithm detected ~50% of all signals*
- *44% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

Compare to high quality subset:

```{r}

diagnose_detection(reference = reference_best, detection = ericks_detection)

```
&nbsp;

<div class="alert alert-info">

- *Erick's algorithm detected ~70% of high quality signals*
- *43% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

# Run ohun energy detector

```{r, eval = FALSE}

ohun_opt_detec <- optimize_energy_detector(reference = reference_best, threshold = c(0.5, 1:4), smooth = 1:4, hold.time = 1:8, min.duration = c(0.05, 0.1, 0.2, 0.3, 0.4, 1, 2), max.duration = c(30, 40, 50), parallel = 10, path = "./data/raw", bp = c(5, 100))

saveRDS(ohun_opt_detec, "./data/processed/high_quality_reference_optimization.RDS")

ohun_opt_detec_all <- optimize_energy_detector(reference = reference_all, threshold = c(0.5, 1:4), smooth = 1:4, hold.time = 1:8, min.duration = c(0.05, 0.1, 0.2, 0.3, 0.4, 1, 2), max.duration = c(30, 40, 50), parallel = 10, path = "./data/raw", bp = c(5, 100))

saveRDS(ohun_opt_detec_all, "./data/processed/all_calls_reference_optimization.RDS")

ohun_opt_detec_ericks <- optimize_energy_detector(reference = ericks_detection, threshold = c(0.5, 1:4), smooth = 1:4, hold.time = 1:8, min.duration = c(0.05, 0.1, 0.2, 0.3, 0.4, 1, 2), max.duration = c(30, 40, 50), parallel = 10, path = "./data/raw", bp = c(5, 100))

saveRDS(ohun_opt_detec_ericks, "./data/processed/ericks_reference_optimization.RDS")

```

# ohun detection performance

Compare to exhaustive reference (only the highest sensitivity iteration is shown)
```{r}

ohun_opt_detec_all <- readRDS("./data/processed/all_calls_reference_optimization.RDS")

ohun_opt_detec_all <- ohun_opt_detec_all[which.max(ohun_opt_detec_all$sensitivity), c(1:11, 17:18)]

ohun_opt_detec_all
# oa_DT <- datatable(ohun_opt_detec_all, editable = list(
#   target = 'row'
# ), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
#   pageLength = 100, autoWidth = TRUE, dom = 'ft'
# ), autoHideNavigation = TRUE, escape = FALSE)
# 
# formatRound(table = oa_DT, columns = sapply(ohun_opt_detec_all, is.numeric), 3)

```
&nbsp;

<div class="alert alert-info">

- *Optimized ohun detection got ~70% of all signals*
- *24% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

Compare to high quality reference (only the highest sensitivity iteration is shown)
```{r}

ohun_opt_detec <- readRDS("./data/processed/high_quality_reference_optimization.RDS")

ohun_opt_detec <- ohun_opt_detec[which.max(ohun_opt_detec$sensitivity), c(1:11, 17:18)]

ohun_opt_detec
# oa_DT <- datatable(ohun_opt_detec, editable = list(
#   target = 'row'
# ), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
#   pageLength = 100, autoWidth = TRUE, dom = 'ft'
# ), autoHideNavigation = TRUE, escape = FALSE)
# 
# formatRound(table = oa_DT, columns = sapply(ohun_opt_detec, is.numeric), 3)

```
&nbsp;

<div class="alert alert-info">

- *Optimized ohun detection got ~74% of high quality signals*
- *32% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

Compare to Erick's detection (only the highest sensitivity iteration is shown)
```{r}

ohun_opt_detec_ericks <- readRDS("./data/processed/ericks_reference_optimization.RDS")

ohun_opt_detec_ericks <- ohun_opt_detec_ericks[which.max(ohun_opt_detec_ericks$sensitivity), c(1:11, 17:18)]

ohun_opt_detec_ericks
# oa_DT <- datatable(ohun_opt_detec_ericks, editable = list(
#   target = 'row'
# ), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
#   pageLength = 100, autoWidth = TRUE, dom = 'ft'
# ), autoHideNavigation = TRUE, escape = FALSE)
# 
# formatRound(table = oa_DT, columns = sapply(ohun_opt_detec_ericks, is.numeric), 3)

```
&nbsp;

<div class="alert alert-info">

- *Optimized ohun detection got ~71% of signals detected by Erick's algorithm*
- *31% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

# Filter using Random Forest

Before random forest filtering
```{r, eval = TRUE}
# best parameters from all signal optimization
ohun_detec <- energy_detector(threshold = 1, smooth = 1, hold.time = 1, min.duration = 0.1, max.duration = 50, parallel = 10, path = "./data/raw", bp = c(5, 100), pb = FALSE)

diagnose_detection(reference_all, ohun_detec)
```

After random forest filtering
```{r, eval = TRUE}
ohun_detec <- label_detection(reference = reference_all, detection = ohun_detec, pb = FALSE)

# measure spectrographic parameters
spectral_parameters <- spectro_analysis(ohun_detec, bp = c(5, 100), fast = TRUE, ovlp = 70, parallel = 10, pb = FALSE)

spectral_parameters$SNR <- sig2noise(ohun_detec, mar = 0.01, pb = FALSE)$SNR

spectral_parameters$class <- ohun_detec$detection.class

spectral_parameters$class[spectral_parameters$class != "false.positive"] <- "true.positive"

# make it a factor for ranger to work 
spectral_parameters$class <- as.factor(spectral_parameters$class)

# run RF model on spectral parameters
  rfm <-
    ranger(
      class ~ .,
      data = spectral_parameters[, !names(spectral_parameters) %in% c("sound.files", "selec")],
      num.trees = 10000,
      importance = "impurity",
      seed = 10
    )

  # table(lab_detec$detection.class)
ohun_detec$pred.class <- rfm$predictions

rf_ohun_detec <- ohun_detec[ohun_detec$pred.class == "true.positive", ]

diagnose_detection(reference = reference_all, detection = rf_ohun_detec, pb = FALSE)

```

&nbsp;

<div class="alert alert-info">

- *Random forest filtered ohun detection got ~68% of all signals*
- *~4% of the detected signals were noise (no bat calls)*

</div>

&nbsp;

---

<font size="4">R session information</font>

```{r session info, echo=F}

sessionInfo()

```