### 2022-07-13 manuscript/manuscript.Rmd

---
title: 'ohun: an R package for optimizing automatic acoustic signal detection'
author:
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
  word_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: yes
bibliography: combined_bibs.bib
citation-style: apa.csl
link-citations: yes
editor_options:
  chunk_output_type: console
---

```{r load packages, echo = FALSE, message = FALSE, warning=FALSE}

# github packages must include user name ("user/package")
# knitr is require for creating html/pdf/word reports
# kableExtra is used to print pretty formatted tables 
# formatR is used for soft-wrapping code
# klippy is used for adding a copy button to each code block

pkgs <- c("kableExtra", "knitr", "rprojroot", "sciRmdTheme", "rmdwc", "whereami", "maRce10/ohun", "ggplot2", "viridis", "rmdwc")

# install/ load packages
out <- lapply(pkgs, function(y) {

  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]

  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  remotes::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  a <- try(require(pkg, character.only = T), silent = T)

  if (!a) remove.packages(pkg)
  })

```

```{r theme setting, echo=FALSE}
# simplest default settings
sciRmdTheme::set.theme()
```

```{r custom data frame printing, echo = FALSE, message=FALSE}

# this is a customized printing style data frames 
# screws up tibble function
tibble <- function(x, ...) { 
  x <- kbl(x, digits=4, align= 'c', row.names = FALSE) 
   x <- kable_styling(x, position ="center", full_width = FALSE,  bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
   asis_output(x)
}

registerS3method("knit_print", "data.frame", tibble)

```

```{r set working directory, echo=FALSE}

# set working directory as project directory or one directory above,
rootdir <- try(rprojroot::find_rstudio_root_file(), silent = TRUE)
if (is(rootdir, "try-error")) rootdir <-  ".."
opts_knit$set(root.dir = rootdir)

```

```{r, fixing citation .bib files, eval = TRUE, echo = FALSE}

# how to cite in Rmarkdown
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

# Here include the path to the libraries for each author. This code will copy/update the library each time the Rmarkdown file is compiled
# Alternatively authors ca
personal_libraries <- c(author_1 = "/home/m/Documents/mendeley_libraries/ohun_paper.bib")

# update bibtex library
for (i in 1:length(personal_libraries))
if (file.exists(personal_libraries[i]))
 file.copy(from = personal_libraries[i], to = file.path(dirname(whereami::thisfile()), paste0(names(personal_libraries)[i], ".bib")), overwrite = TRUE)

## combine .bib files in compiled.bib ###
# list bib files
bibs <- list.files(pattern = ".bib$", path = dirname(whereami::thisfile()), full.names = TRUE)

# remove combined from pool
bibs <- bibs[basename(bibs) != "combined_bibs.bib"]

#  read in a list
if (length(bibs) > 0){
  combined_bibs_l <- lapply(bibs, readLines)
  names(combined_bibs_l) <- gsub("\\.bib$", "", basename(bibs))

# combine bibs in a single one
combined_bibs <- unlist(combined_bibs_l)

# save pool libraries as combined_bibs.bib
writeLines(text = combined_bibs, file.path(dirname(whereami::thisfile()), "combined_bibs.bib"))

# check if there are duplicated keys
keys <- grep("@article{", combined_bibs, fixed = TRUE, value = TRUE)
keys <- gsub("@article{", "", keys, fixed = TRUE)
keys <- gsub(",", "", keys, fixed = TRUE)

tab_keys <- table(keys)

if (anyDuplicated(keys)){
  print(paste0(sum(tab_keys > 1), " duplicate(s) references found in combined_bibs.bib"))
}
} else writeLines(text = "empty", file.path(dirname(whereami::thisfile()), "combined_bibs.bib"))

```

```{r read data, echo=FALSE}

path_zebra_finch <- file.path(rootdir, "data/raw/taeniopygia/")

# read data
manual_ref_thy <- read.csv(file.path(rootdir, "//data/raw/thyroptera/", "manual_annotations_thyroptera.csv"))

manual_ref_tae <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "manual_selections_Taeniopygia.csv"))

```

Marcelo Araya-Salas ^1,^ ^2,^ ^3^ *,
Grace Smith-Vidaurre ^4^,
Gloriana Chaverri ^3^,
Fabiola Chirino ^1^ &
Alejandro Rico-Guevara ^5^

&nbsp;

^1^ Centro de Investigación en Neurociencias, Universidad de Costa Rica, San José, Costa Rica

^2^ Escuela de Biología, Universidad de Costa Rica, San José, Costa Rica

^3^ Sede del Sur, Universidad de Costa Rica, Golfito, Costa Rica

^4^ GRACE'S AFFILIATION

^5^  Department of Biology, University of Washington, Seattle, USA

\* *To whom correspondence should be addressed*

### Keywords:

---

&nbsp;

```{r word count, eval = TRUE, echo = FALSE, results='asis'}

# word count without code chunks
word_count <- rmdcount(whereami::thisfile())

# word count with code chunks
word_count_chunk <- rmdcount(whereami::thisfile(), exclude='')

# print results
cat(paste0("Word count: ", word_count[3], " (including code chunks: ", word_count_chunk[3], ")"))

```

Check [this link](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html) for details on how to add citations/bibliography in Rmarkdown

---

## Abstract

Animal acoustic signals are widely used in a diversity of research areas. This is partly due to the relative ease in which they can be registered for wide range of taxonomic groups and research settings. However, bioacoustic research can easily generate large amounts of data which might proof challenging to analyze in a timely manner. Many tools for the automatic detection of signals are currently available. However, choosing the right approach for a specific task might turn difficult as there are no available tools that can provide a common framework for evaluating detection performance. Here we present the new R package ohun. The package is intended to facilitate the automatic detection of acoustic signals, by providing functions to diagnose and optimize detection routines as well as compare the performance of different detection approaches. The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines, using common signal detection theory indices. This can be done with routine outputs imported from other software and from detection run within the package itself. The package also provides a set functions to organize acoustic data sets in a format amenable for detection analyses. It also includes an implementation of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection. We included examples on how to evalute detections from external software and showcase the package usage with study cases on zebra-finch (*Taenopygia gutata*) songs and Spix's disc-winged bat (*Thyroptera tricolor*) ultrasonic social calls. Finally, we provide some general suggestions to improve detection performance.

## Introduction

Animal acoustic signals are widely used to investigate a variety of questions in highly diverse areas, ranging from neurobiology to community ecology and evolutionary biology. The wide usage of animal sounds in research is partly due to the fact that they can be easily registered using non-intrusive methods, can be obtained in a variety of settings from laboratories to natural areas, and the equipment required for registering an analyzing these signals has become increasingly inexpensive. In addition, the existence of online repositories and growing number has facilitated the study of these traits at larger taxonomic and geographic scales. However, the adoption of bioacustic approaches in research can also imply large amounts of data (i.e. lots of recordings), which can be challenge to analyzed manually by human observers. As a result, a growing number of computational tools for the analysis of acoustic features on those signals is increasingly available (reviewd by @Stowell2022), reflecting the need of automated approaches for efficiently conducting these analyses.

<!-- Many of these tools are supplied in the form of freely available software written on programming environments familiar to the scientific community (seewave, warbleR, scikit-mad), further simplifying their implementation for research. -->
<!-- Bioacoustic research can easily generate large amounts of data, which sometimes may proof challenging to analyze in a timely manner. Annotations that require precise location of signals in frequency and time, most commonly used in behavioral and phenotypic evolution research, remain among the most time consuming phases of analysis for this type of data. Fortunately, several tools have been develop to assist researchers in this matter. Indeed, new methods for automatic annotation of acoustic signals are continuously developed . -->

The growing availability of tools for automatic detection of acoustic events, particularly as free software, is expected to further simplify acoustic data processing, making it accessible to a wider range of users and scientific questions. However, this diversity of tools also posits a challenge as it can be hard to navigate. In this regard, the use of standard approaches for evaluating the performance of automatic detection tools might prove helpful to inform the decision about which method better fit our study system. The performance of automatic acoustic signal detection routines have been evaluated using standard indices from signal detection theory (CITATIONS). In its basic form, performance is assessed by comparing the output of a detection routine against a "gold standard" reference in which all the target signals have been annotated. This comparison allows to quantify the number of correctly detected signals (true positives), wrongly detected signals (false positives) and missed signals (false negatives) as well as additional metrics derived from these indices (e.g. recall, precision). However, the fact that acoustic signals are not being evaluated as discrete classification units (as opposed to for instance identifying species from pictures) demand additional information to fully diagnose detection performance. This is particularly relevant when the precise time position of signals is needed as is common for research in which measures of the acoustic structure of signals is the main goal. For instance, the same signal can be detected as several separated signals, the inferred time position can be far off from the the target signal position, or several signals can be detected as a single one. Therefore, metrics that account for this other performance dimensions can be a valuable tool to properly diagnose automatic acoustic signal detection.

Here we present the new R package ohun. The package is intended to facilitate the automatic detection of acoustic signals, providing functions to diagnose specific aspects of acoustic detection routines in order to simplify their optimization. The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines. This can be done with routine outputs imported from other software as well as detection run within the package itself. The package also provides a set functions to explore acoustic data sets and organize them in a format amenable for detection analyses. In addition, it offers implementations of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection. We explain how to explore and format acoustic data sets, how acoustic signal detection routines can be evaluate and then showcase the package usage with study cases on zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bats (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating signals and ultrasonic social calls).

## Formatting acoustic data sets

Having some sense of the format and size of the sound files to be analyzed is an important step for avoiding downstream errors and informing our time performance expectations. Several functions can facilitate getting acoustic data sets in the right format prior to automatic detection.  The function `feature_acoustic_data` prints a summary of the duration, size and format of all the recordings in a folder. Here we explore the acoustic data set of Zebra finch's songs (suppl. mat):

```{r, eval = FALSE, echo = TRUE}

# working directory
path_zebra_finch <- "path_to_zebra_finch_files"

feature_acoustic_data(path = path_zebra_finch)

```

```{r, eval = TRUE, echo = FALSE}

feature_acoustic_data(path = path_zebra_finch)

```

In this case all recordings have format (.wav files, 44.1 kHz sampling rate, 16 bit resolution and a single channel). We can also check the duration and size of files. This information is important as some tuning parameters of detection routines can behave differently depending on file format (e.g. time window size is affected by sampling rate) or simply because some software might can only work on specific sound file formats. In addition, long sound files might be hard to analyzed in regular computers and might have to be split in shorter clips. In the latter case the function `split_acoustic_data` can be used to produce those clips:

```{r, eval = FALSE, echo = FALSE}
source("~/Dropbox/R_package_testing/ohun/R/split_acoustic_data.R")
split_info <- split_acoustic_data(path = path_zebra_finch, sgmt.dur = 5)

write.csv(split_info, "./data/processed/split_data_for_manuscript.csv", row.names = FALSE)

```

```{r, eval = FALSE, echo = TRUE}

split_info <- split_acoustic_data(path = path_zebra_finch, sgmt.dur = 5)

head(split_info)

```

```{r, eval = TRUE, echo=FALSE}

split_info <- read.csv(file.path(rootdir, "data/processed/split_data_for_manuscript.csv"))

head(split_info)
```

The output shows to which time segments in the original sound files the clips belong to. If an annotation table is supplied (argument 'X') the function will adjust the annotations so they refer to the position of the signals in the clips. This can be helpful when reference tables have been annotated on the original sound files.

Annotations can also be explored using the function `feature_reference`. It returns about the mean and range of signal and gap duration (time intervals between selections), bottom and top frequency and of the number of annotations by sound file. If the path to the sound files is supplied the duty cycle (fraction of a sound file corresponding to signals) and peak amplitude (highest amplitude in a detection) are also returned:

```{r, eval = FALSE, echo = TRUE}

# read reference annotations
manual_ref_tae <- read.csv(file.path(path_zebra_finch, "manual_selections_Taeniopygia.csv"))

# explore annotations
feature_reference(reference = manual_ref_tae, path = path_zebra_finch)

```

```{r, eval = TRUE, echo = FALSE}

manual_ref_tae <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "manual_selections_Taeniopygia.csv"))

feature_reference(reference = manual_ref_tae, path = path_zebra_finch)

```

## Diagnosing detection performance

The package makes use of signal detection theory indices to evaluate detection performance. Signal detection theory deals with the process of recovering signals (i.e. target signals) from background noise (not necessarily acoustic noise) and it’s widely used for optimizing this decision making process in the presence of uncertainty. During a detection routine, the detected ‘items’ can be classified into 4 classes: true positives (TPs, target signals correctly identified as signal), false positives (FPs, noise incorrectly identified as ‘signal’), false negatives (FNs, signals incorrectly identified as noise) and true negatives (TNs, background noise correctly identified as noise). However, TNs cannot always be easily defined in the context of acoustic signal detection as noise cannot always be partitioned in discrete units. Hence, the package makes use of TPs, FPs and FNs to calculate three additional indices that can further assist with evaluating the performance of a detection routine.

 - Recall: correct detections relative to total detections (a.k.a. true positive rate or sensitivity; TPs / (TPs + FNs))
- Precision: correct detections relative to total detections (TPs / (TPs + FPs)).
- F1 score: combines recall and precision as the harmonic mean of these two, providing a single value for evaluating performance (a.k.a. F-measure or Dice similarity coefficient).

A perfect detection will have no false positives or false negatives, which will result in both recall and precision equal to 1. However, perfect detection cannot always be reached and some compromise between detecting all target signals plus some noise (recall = 1 & precision < 1) and detecting only target signals but not all of them (recall < 1 & precision = 1) might be warranted. The right balance between these two extremes will be given by the relative costs of missing signals and mistaking noise for signals, given the specific goals of the study. These indices provide an useful framework for diagnosing and optimizing the performance of a detection routine.

The package offers tools to evaluate the performance of an acoustic signal detection based on the indices described above. To accomplish this, annotations derived from a detection routine are compared against a reference table containing the time position of all target signals in the sound files. For instance, the following code evaluates a routine run in Raven Pro 1.6 (XXXX) using the "band limited energy detector" option (minimum frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.54989 s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a subset of the Zebra finch recordings described below:

```{r, eval = TRUE, echo = FALSE}

library(Rraven)

raven_detec <- imp_raven(path = file.path(rootdir, "/data/raw/"), files = "Raven.Band.Limited.Energy.Detector.selections.txt", warbler.format = TRUE)

# raven_ref <- manual_ref[manual_ref$sound.files %in% raven_detec$sound.files, ]

```

```{r, eval = TRUE, echo = TRUE}

diagnose_detection(reference = manual_ref_tae, detection = raven_detec)

```

The output shows the indices described above. The function also allows to detail those indices by sound file. Here we show the first 6 files:

```{r, eval = TRUE, echo = TRUE}

diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by.sound.file = TRUE)

head(diag_raven)
```

Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize a detection. The process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by ohun, which we depict in the following section. Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources rather than a direct comparison of performance to ohun.

## Signal detection with ohun

The package offers two methods for automatic signal detection: template-based and energy-based detection. These methods are better suited for highly stereotyped or good signal-to-noise ratio (SNR) signals, respectively. If the target signals do not fit these requirements, more elaborated methods (i.e. machine/deep learning approaches) are warranted.

## Study cases

### Template detection on ultrasonic social calls of  Spix's disc-winged bats

We recorded `r length(unique(manual_ref_thy$sound.files))` individuals of Spix's disc-winged bats at Baru Biological Station, in south-western Costa Rica in January 2020. Bats were captured at their roosting sites (furled leaves of Zingeberaceae plants). Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5 min period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm plugged into a laptop computer running Avisoft-Recorder software. Recordings were made at a sampling rate of 500 kHz and an amplitude resolution of 16 bits.

Recordings were manually annotated using Raven Pro 1.6 (XXXX). Annotations were created by visual inspection of spectrograms, in which the start and end of signals were determined by the location of the continuous traces of power spectral entropy of the target signals. A total of `r nrow(manual_ref_thy)` calls were annotated (~`r round(nrow(manual_ref_thy) / 30, 0)` calls per recording). Annotations were made with a time window of 200 samples and 70% of overlap. Annotations were then imported into R using the package Rraven (XXXX).

Inquiry calls of Spix's disc-winged bats are structurally stereotyped. Most variation is found between individuals although the basic form of a short, downward broadband frequency modulation is always shared (Fig. BAT-SPECTRO, Araya-Salas et al 2021 ontogeny).

```{r catalog, out.width = "100%", echo = FALSE, fig.align= "left",fig.cap="Fig. BAT-SPECTRO. Example spectrograms of Spix's disc-winged social calls for each of the 30 recordings used in the analysis. The highest signal-to-noise ratio call by sound file are shown. The time scale range is 71 ms and the frequency range 10-44 kHz", error=TRUE}

# knitr::include_graphics("~/Dropbox/Projects/ohun_paper/data/raw/thyroptera/Catalog_p1.jpeg")
knitr::include_graphics((file.path(rootdir, "data/raw/thyroptera/Catalog_p1.jpeg")))
```

Template-based detection is a useful approach when there is little structural differences in the target signals. We used this approach in ohun to detect inquiry calls. To do this, we tested the performance of three acoustic templates on a training subset of five sound files. The function `get_templates` finds several signals representative of the variation in signal structure. This function measures several spectrographic parameters which are then summarized using Principal Component Analysis. The first two components are used to project the acoustic space. On this space the function defines sub-spaces as equal-size slices of a circle centered at the centroid of the acoustic space. Templates are then selected as those closer to the centroid within each of the sub-spaces, including the centroid for the entire acoustic space. The user needs to define the number of sub-spaces in which the acoustic space will be split.

```{r, eval = FALSE}

# read manual annotations
manual_ref_thy <- read.csv("manual_annotations_thyroptera.csv")

# get random subset of 5 sound files for training
set.seed(1)
train_files <- sample(unique(manual_ref_thy$sound.files), size = 5)
train_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% train_files, ]

# the rest for testin g
test_files <- setdiff(manual_ref_thy$sound.files, train_files)
test_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% test_files, ]

# find templates
templates <- get_templates(train_ref, path = data_path,  bp = c(10, 50), ovlp = 70, wl = 200, n.sub.spaces = 3)

```

```{r, echo = FALSE, fig.height= 6, fig.cap= "FIG. ACOUSTIC-SPACE. Acoustic space defined as the the first two component of a Principal Component Analysis on spectrographic parameters. Templates are selected as those closer to the centroid for each of the sub-spaces. Gray dashed lines delimit the region of sub-spaces. Yellow circles around points highlight the position of the signals selected as templates."}

# read manual annotations
manual_ref_thy <- read.csv(file.path(rootdir, "/data/raw/thyroptera/", "manual_annotations_thyroptera.csv"))

# get random subset of 5 sound files for training
set.seed(1)
train_files <- sample(unique(manual_ref_thy$sound.files), size = 5)
train_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% train_files, ]

# the rest for testing
test_files <- setdiff(manual_ref_thy$sound.files, train_files)
test_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% test_files, ]

source("~/Dropbox/R_package_testing/ohun/R/get_templates.R")

templates <- get_templates(train_ref, path = file.path(rootdir, "/data/raw/thyroptera/"),  bp = c(10, 50), ovlp = 70, wl = 200, n.sub.spaces = 3)

```

```{r, eval = TRUE, echo = FALSE}

data_path <- file.path(rootdir, "/data/raw/thyroptera/")

manual_ref_thy <- read.csv(file.path(data_path, "manual_annotations_thyroptera.csv"))

# get random subset of 5 sound files for training
set.seed(1)
train_files <- sample(unique(manual_ref_thy$sound.files), size = 5)
train_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% train_files, ]

# the rest for testing
test_files <- setdiff(manual_ref_thy$sound.files, train_files)
test_ref <- manual_ref_thy[manual_ref_thy$sound.files %in% test_files, ]
```

The output of the `get_templates` includes an acoustic space plot (FIG. ACOUSTIC-SPACE) in which the position of the signals selected as templates is highlighted. In the following code we used those templates for detecting calls. The code iterates a template-based detection on the training data set across a range of correlation thresholds, in order to find the combination of threshold and template with the best performance.

```{r, eval = FALSE}

# get correlation vectors
corr_templ_train <- template_correlator(
  templates = templates,
    path = data_path, 
  files = unique(train_ref$sound.files), 
  hop.size = 10, 
  ovlp = 70
  )

# evaluate detection for different correlation thresholds
opt_detec_train <- optimize_template_detector(
  reference = train_ref, 
  template.correlations = corr_templ_train,
  threshold = seq(0.05, 0.5, 0.01)
  )

```

Note that the correlation vectors are estimated first (i.e. vectors of correlation values across sound files, `template_correlator()`) and then the correlation thresholds are optimized on them (`optimize_template_detector()`). The output of `optimize_template_detector()` contains the detection performance indices for each combination of templates and thresholds. Table TEMPLATE PERFORMANCE shows the two highest performance runs for each template.

```{r, eval = TRUE, echo = FALSE}

opt_detec_train <- readRDS(file.path(rootdir, "/data//processed/optimization_results_3_templates_thyroptera.RDS"))

# opt_detec_train <- opt_detec_train[order(opt_detec_train$templates, opt_detec_train$f1.score, decreasing = TRUE), ]

# sub_opt_detec_train <- opt_detec_train[ave(-opt_detec_train$f1.score, opt_detec_train$templates, FUN = rank) <= 4, ]

sub_odt <- lapply(unique(opt_detec_train$templates), function(x) {

    X <- opt_detec_train[opt_detec_train$templates == x, ]
    X <- X[order(X$f1.score, decreasing = TRUE), ]
    return(X[1:2, ])
    })

sub_opt_detec_train <- do.call(rbind, sub_odt)

kbl <- kable(sub_opt_detec_train[, c("threshold", "templates", "true.positives", "false.positives", "false.negatives", "recall", "precision", "f1.score")], row.names = FALSE, caption = "TABLE TEMPLATE PERFORMANCE. Performance diagnostic of template-based detections using four templates across several threshold values. Only the two highest performance iterations for each template are shown.")

kable_styling(kbl, bootstrap_options = c("striped", "hover",  full_width = FALSE))

```

We can explore the performance of each template in more detail by looking at the change in F1 score across thresholds (FIG. THRESHOLD vs F1.SCORE).

```{r, eval = TRUE, echo = FALSE, fig.height= 3, fig.align='center', fig.cap= "FIG. THRESHOLD vs F1.SCORE. Shows the changes in F1 score across the range of threshold values"}

agg_f1 <- aggregate(f1.score ~ threshold + templates, data = opt_detec_train,
mean)

ggplot(agg_f1, aes(x = threshold, y = f1.score, group = templates, color = templates)) +
    geom_line() + 
    geom_point() + 
    scale_color_viridis_d(end = 1, labels = c("Centroid", "1st", "2nd", "3rd"), alpha = 0.7) + 
    labs(x= "Cross-correlation threshold", y = "F1 score", color = "Templates")

```

In this example the "centroid" template, produced the best performance (TABLE TEMPLATE PERFORMANCE; FIG. THRESHOLD vs F1.SCORE). Hence, we will use this template for detecting calls on the rest of the data. The following code extracts this template from the reference annotation table and use it to find inquiry calls on the testing data set:

```{r, detect on all data 67, eval = FALSE, echo = TRUE}

# get correlation vectors for test files
corr_templ_test <- template_correlator(
  templates = templates[templates$sound.file == "centroid", ],
    path = data_path, files = unique(test_ref$sound.files), 
  hop.size = 10, 
  ovlp = 70
  )

# detect on test files
detec_test <- template_detector(
  template.correlations = corr_templ_test,
    threshold = 0.45
  )

diagnose_detection(reference = test_ref, detection = detec_test)

```

```{r, eval = TRUE, echo = FALSE}

detec_test <- readRDS(file.path(rootdir, "/data/processed/detection_thyroptera_all_67%_template.RDS"))

diagnose_detection(reference = test_ref, detection = detec_test)
```

The last line of codes evalutes the detection on the test data set, which shows a good performance for both recall and precision.

### Energy-based detection on zebra finch songs

We used recordings from 18 zebra finch males. These recordings were obtained ....

Zebra finch vocalizations are composed by multiple elements (i.e. distinct patterns of continuous traces of power spectral entropy in the spectrogram separated by time gaps) that can vary importantly in key features as duration and frequency range (ZEBRAFINCH-SPECTRO). However, as the recorded signals show a good signal-to-noise ratio, they can potentially be detected using an energy-based approach.  
```{r catalog zebra f, out.width = "100%", echo = FALSE, fig.align= "left",fig.cap="Fig. ZEBRAFINCH-SPECTRO. Example spectrograms of male zebra finch songs for each of the 18 sound files used in the analysis. The highest signal-to-noise ratio call by sound file are shown. The time scale range is 359 ms and the frequency range 0-11 kHz. Signals have been highlighted for visualization purposes only.", error=TRUE}

knitr::include_graphics((file.path(rootdir, "data/raw/taeniopygia/same.time-Catalog_p1.jpeg")))

```

The following code loads the reference annotations and split them in two data set for training (3 sound files) and testing (15 sound files):
```{r, eval = TRUE, echo = FALSE}

manual_ref_tae <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "manual_selections_Taeniopygia.csv"))

set.seed(450)    
train_files <- sample(unique(manual_ref_tae$sound.files), 3)

test_files <- setdiff(manual_ref_tae$sound.files, train_files)

train_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% train_files, ]
test_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% test_files, ]

```

```{r, eval = FALSE, echo = TRUE}

manual_ref_tae <- read.csv("manual_selections_Taeniopygia.csv")

set.seed(450)    
train_files <- sample(unique(manual_ref_tae$sound.files), 3)    
test_files <- setdiff(manual_ref_tae$sound.files, train_files)

train_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% train_files, ]
test_ref <- manual_ref_tae[manual_ref_tae$sound.files %in% test_files, ]

```

Now we can optimize the detection parameters using the function `optimize_energy_detector`. This function runs a detection for all possible combinations of tuning parameters. The code below tries three minimum duration and maximum duration values and two hold time values (which merges signals within the specified time interval):

```{r}
feature_reference(test_ref)
```

```{r  optimize automated detection, eval = FALSE, echo = TRUE}

opt_det_train <- optimize_energy_detector(
  reference = train_ref, 
  files = train_files, 
  threshold = c(1, 5), 
  hop.size = 11.6, 
  smooth = c(5, 10), 
  hold.time = c(0, 5), 
  min.duration = c(5, 15, 25), 
  max.duration = c(275, 300, 325), 
  bp = c(0.5, 10)
)

```

The output (`opt_det_train`) shows the performance indices for each of those combinations, here we show the 10 combinations with the highest F1 score:

```{r, eval = FALSE, echo = TRUE}

# subset with highest performance
opt_det_train <- opt_det_train[order(opt_det_train$f1.score, decreasing = TRUE), ]

head(opt_det_train, 10)
```

```{r, eval = TRUE, echo = FALSE}

opt_det_train <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "detection_optimization.csv"))

# subset with highest performance
sub_opt_det_train <- opt_det_train[order(opt_det_train$f1.score, decreasing = TRUE), ]

sub_opt_det_train <- sub_opt_det_train[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f1.score")]

sub_opt_det_train
```

We now can use the tuning parameter values that produce the best performance to detect signals on the test data set:

```{r, select best params, eval = FALSE, echo = TRUE}

best_param <- opt_det_train[which.max(opt_det_train$f1.score), ]

det_test <- energy_detector(
  files = test_files, 
  threshold = best_param$threshold, 
  hop.size = 11.6, 
  smooth = best_param$smooth, 
  hold.time = best_param$hold.time, 
  min.duration = best_param$min.duration, 
  max.duration = best_param$max.duration, 
  bp = c(0.5, 10)
)
```

As our reference annotations include all sound files we can evaluate the performance of the detection on the test set as well:

```{r , eval = TRUE, echo = FALSE}

det_test <- read.csv(file.path(rootdir, "/data/processed/taeniopygia", "detection_test_data.csv"))

```

```{r, diagnose, eval = TRUE, echo = TRUE}

diagnose_detection(reference = test_ref, detection = det_test, by.sound.file = FALSE)

```
&nbsp;

The performance on the test data set was also acceptable with a F1 score of 0.95. Note that in the example we use a small subset of sound files for training. More training data might be needed for optimizing a detection routine on larger data sets or recordings with more variable signals or background noise levels.

### Additional tools

The package offers additional tools to simplify acoustic signal detection. The function 'feature_reference' allows to extract information about the duration and frequency range of acoustic signals than can then be used to define tuning parameter values. This can be particularly useful for energy-based detection in which time and frequency attributes of the target signals are required. Detected signals can be labeled as false or true positives with the function 'label_detection'. This allows users to explore the structure of false positives and figure out ways to exclude them. The function 'filter_detections' can be used to remove ambiguous signals (i.e. those labeled as split or merged detection) keeping only those that maximize a specific criterium (i.e. the highest template correlacion). Finally, note that  several templates representing the range of variation in signal structure can be used to detect semi-stereotyped signals when running template-based detection ('template_detection' function).

## Discusion

Here we have showed how to evaluate the performance of acoustic signal detection routines using the package ohun. The package can evaluate detection outputs imported from other software, as well as its own detection routines. The latter can be iterated over combinations of tuning parameters in order to find those values that optimize a detection. Despite signal detection indices being commonly reported when presenting new automatic detection methods, to our knowledge widely applicable performance evaluating routines have not been made available in a free, open source platform. Providing a common framework for acoustic signal detection evaluation can simplify comparing performances of different tools and the selection of those better suited to our system and research question. Note that the tools offered by ohun for diagnosing a detection should not be necessarily limited to acoustic data. Any type of automatic detection in which the time of occurrence of discrete events needs to be found can be evaluated and optimized by comparing it to a reference annotation. For instance detection of specific behaviors in video analysis of animal motor activity (Sturman et al 2020; Hsu & Yttri 2021; DeepEthogram).

The package provides two native detection methods: template-based  and energy-based detection. Compared to new deep learning approaches for finding the occurrence of acoustic signals, the two native methods are relatively simple tools. However, these methods have been widely used by the bioacoustic community and, under the appropriate conditions can reach adequate performance, as we have shown in our two study cases. Deep learning methods tend to require larger computational power and more complex training routines. This might bring unnecessary difficulties when dealing when less challenging detection tasks. Therefore the availability of a wide range of approaches is desired in order to simplify finding the most appropriate tool for the intricacies of our study system and research goals. Note that the tools offered in ohun can be used in a pipeline in which detected signals are further classified using more elaborated discrimination algortihms. Acoustic parameters can be used to quantify the structure of the signals and tell apart target from non-target signals.

Detection routines can take a long time when working with large amounts of acoustic data (e.g. large recordings and/or many files). These are some tips that can help make a routine more time-efficient. 1. Always test procedures on small data subsets. Make sure you are getting decent results on a small subset of recordings before trying to scale up the analysis. 2. Template-based detection is almost always faster than energy-based detection. 4. Run routines in parallel. Parallelization (i.e. the ability to distribute tasks over several cores in your computer) can significantly speed-up routines. All functions for automatic detection and performance evaluation allow users to run analysis in parallel (see `parallel` argument in those functions). Hence a computer with several cores can be helpful for improving efficiency. 4. Sampling rate matters. Detecting signals on low sampling rate files is faster, so we must avoid having Nyquist frequencies much higher than the highest frequency of the target signals. 5. Try using a computer with lots of RAM memory or a computer cluster for working on large amounts of data. These rules are not restricted to ohun and can also be helpful to speed-up routines in other software packages.

There are some additional things to be considered when aiming to automatically detect acoustic signals. When running automatic signals detection try to use your knowledge about the signal structure to determine the initial range. This can be extremely helpful for narrowing down possible parameter values, particularly for energy-based detection. As a general rule, if people have a hard time figuring out where a target signal occurs, detection algorithms will also have a hard time. In cases in which occurrences are ambiguous low performances are expected. Make sure reference tables contain all target signals and only the target signals, otherwise performance optimization can be misleading as the performance of a detection cannot be better than the reference itself. Lastly, avoid having overlapping signals or several signals as a single one (like a multi-syllable vocalization) in the reference table when running an energy-based detector, as they are likely to be identified as separeted units.

https://cran.r-project.org/web/packages/bioacoustics/vignettes/tutorial.html

## References

Sturman, O., von Ziegler, L., Schläppi, C. et al. Deep learning-based behavioral analysis reaches human accuracy and is capable of outperforming commercial solutions. Neuropsychopharmacol. 45, 1942–1952 (2020). https://doi.org/10.1038

Hsu, A.I., Yttri, E.A. B-SOiD, an open-source unsupervised algorithm for identification and fast prediction of behaviors. Nat Commun 12, 5188 (2021). https://doi.org/10.1038/s41467-021-25420-x