### 2022-07-13 README.Rmd
Update README.Rmd

---
title: ohun_paper_analyses
author: Marcelo Araya Salas
output: github_document
date: `r format(Sys.Date(), "%d-%m-%Y")`
editor_options:
  chunk_output_type: console
---

<!-- Short Description  -->
Data analysis and manuscript writting for method paper on the R package ohun

*Updated on `r Sys.time()`*

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}

library(rmarkdown)

knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%"
)
```

## Table of contents
* [General info](#general-info)
    * [To do list](#to-do-list)
      * [Contact](#contact)

## General info

Data analysis and manuscript writting for method paper on the R package ohun

## In progress

*  Writting

## To-do list

* Add citations

## Status
Project is: _in progress_

## Contact
Created by [Marcelo Araya-Salas](marceloarayasalas.weebly.com/)