### 2023-05-24 manuscript/manuscript.tex
many changes

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={ohun: an R package for diagnosing and optimizing automated sound event detection},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{ohun: an R package for diagnosing and optimizing automated sound
event detection}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{3}
\tableofcontents
}
Marcelo Araya-Salas \textsuperscript{1,} \textsuperscript{2,}
\textsuperscript{3} *, Grace Smith-Vidaurre \textsuperscript{4,}
\textsuperscript{5,} \textsuperscript{6}, Gloriana Chaverri
\textsuperscript{3}, \textsuperscript{7} Juan Carlos Brenes
\textsuperscript{1} Fabiola Chirino \textsuperscript{1} \& Alejandro
Rico-Guevara \textsuperscript{8,} \textsuperscript{9}

~

\textsuperscript{1} Centro de Investigación en Neurociencias,
Universidad de Costa Rica, San José, Costa Rica

\textsuperscript{2} Escuela de Biología, Universidad de Costa Rica, San
José, Costa Rica

\textsuperscript{3} Sede del Sur, Universidad de Costa Rica, Golfito,
Costa Rica

\textsuperscript{4} Laboratory of Neurogenetics of Language, Rockefeller
University, New York, NY, USA

\textsuperscript{5} Rockefeller University Field Center, Millbrook, NY,
USA

\textsuperscript{6} Department of Biological Sciences, University of
Cincinnati, Cincinnati, OH, USA

\textsuperscript{7} Smithsonian Tropical Research Institute, Panama
City, Panamá

\textsuperscript{8} Department of Biology, University of Washington,
Seattle, USA

\textsuperscript{9} Burke Museum of Natural History and Culture,
University of Washington, Seattle, USA

* \emph{To whom correspondence should be addressed}

\textbf{Keywords}: bioacoustics, automatic detection, animal
vocalizations

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

~

Word count: 4279 (including code chunks: 5925)

Check
\href{https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html}{this
link} for details on how to add citations/bibliography in Rmarkdown

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Animal acoustic signals are widely used in diverse research areas due to
the relative ease with which sounds can be registered across a wide
range of taxonomic groups and research settings. However, bioacoustics
research can quickly generate large data sets, which might prove
challenging to analyze promptly. Although many tools are available for
the automated detection of sounds, choosing the right approach can be
difficult only a few tools provide a framework for evaluating detection
performance. Here we present ohun, an R package intended to facilitate
automated sound detection. ohun provides functions to diagnose and
optimize detection routines and compare performance among different
detection approaches. The package uses reference annotations containing
the time position of target sounds in a training data set to evaluate
detection routines performance using common signal detection theory
indices. This can be done both with routine outputs imported from other
software and detections run within the package. The package also
provides functions to organize acoustic data sets in a format amenable
to detection analyses. ohun also includes energy-based and
template-based detection methods --two commonly used automatic
approaches in bioacoustic research. We show how ohun automatically can
be used to detect vocal signals with case studies of adult male zebra
finch (Taenopygia gutata) songs and Spix's disc-winged bat (Thyroptera
tricolor) ultrasonic social calls. We also include examples of how to
evaluate the detection performance of ohun and external software.
Finally, we provide some general suggestions to improve detection
performance.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Animal acoustic signals are widely used to address a variety of
questions in highly diverse areas, ranging from neurobiology
(\protect\hyperlink{ref-Burgdorf2011}{Burgdorf \emph{et al.} 2011};
\protect\hyperlink{ref-schoneich2020}{S 2020}) to taxonomy
(\protect\hyperlink{ref-Kohler2017}{Köhler \emph{et al.} 2017};
\protect\hyperlink{ref-gew2019}{Gwee \emph{et al.} 2019}), community
ecology (\protect\hyperlink{ref-zsebok2021}{Zsebők \emph{et al.} 2021};
\protect\hyperlink{ref-tiwari2022}{Tiwari \& Diwakar 2022}) and
evolutionary biology
(\protect\hyperlink{ref-Medina-Garcia2015a}{Medina-García \emph{et al.}
2015}; \protect\hyperlink{ref-Odom2021}{Odom \emph{et al.} 2021}). The
profuse usage of animal sounds in research is partly due to the fact
that they can be easily collected using non-invasive methods. Animal
sounds can also be obtained in a variety of settings from laboratories
to natural areas, while the equipment required for collecting an
analyzing these sounds has become increasingly inexpensive and broadly
accesible (\protect\hyperlink{ref-Blumstein2011}{Blumstein \emph{et al.}
2011}; \protect\hyperlink{ref-sugai2019}{Sugai \emph{et al.} 2019}). In
addition, the existence of online repositories has facilitated the study
of these communication signals at larger taxonomic and geographic
scales. However, the adoption of bioacustic approaches in research may
also imply large amounts of data (i.e.~lots of recordings), which can be
challenging to analyzed manually (\protect\hyperlink{ref-gibb2019}{Gibb
\emph{et al.} 2019}). As a result, a growing number of computational
tools for the automatic detection of animal sounds on is increasingly
available (reviewed by \protect\hyperlink{ref-Stowell2022}{Stowell
2022}), reflecting the need of automated approaches for efficiently
conducting these analyses (\protect\hyperlink{ref-gibb2019}{Gibb
\emph{et al.} 2019}).

The growing availability of tools for automatic detection of acoustic
events, particularly as free software, is expected to further simplify
acoustic data processing, making it accessible to a wider range of users
and scientific questions. However, this diversity of automated detection
tools also posits a challenge as it can be difficult to navigate
(\protect\hyperlink{ref-Stowell2022}{Stowell 2022}). In this regard, the
use of standard approaches for evaluating the performance of automatic
detection tools might prove helpful to inform researchers' decisions
about which method better fits a given question and study system
(\protect\hyperlink{ref-knight2017}{Knight \emph{et al.} 2017}). The
performance of automated sound event detection routines have typically
been evaluated using standard indices from signal detection theory
(\protect\hyperlink{ref-macmillan2004}{Macmillan \& Creelman 2004};
\protect\hyperlink{ref-knight2017}{Knight \emph{et al.} 2017};
\protect\hyperlink{ref-balantic2020}{Balantic \& Donovan 2020}). In its
basic form, performance is assessed by comparing the output of a
detection routine against a `gold standard' reference in which all the
target sounds have been annotated (hereafter called `reference
annotation'). This comparison facilitates quantifying the number of
correctly detected sounds (true positives), wrongly detected sounds
(false positives) and missed sounds (false negatives) as well as
additional metrics derived from these indices (e.g.~recall, precision).
However, the fact that sound events are not being evaluated as discrete
classification units (as opposed to for instance identifying species
from pictures), but are rather embedded within a continuous string of
sound, demand additional information to fully diagnose the temporal
precision detection performance. This is particularly relevant when the
precise time position of sounds is needed as is common when measuring
the acoustic structure of sounds is the main goal
(\protect\hyperlink{ref-Araya-Salas2017f}{Araya-Salas \& Smith-Vidaurre
2017}). For instance, the same signal can be detected as several
separated sounds, the inferred time position can be offset from the the
target signal position, or several sounds can be detected as a single
one. Therefore, metrics that account for these additional performance
dimensions are a valuable tool to properly diagnose automatic sound
event detection.

Here we present the new R package \emph{ohun}. This package is intended
to facilitate the automatic detection of sound events, providing
functions to diagnose particular aspects of acoustic detection routines
in order to simplify their optimization. The package makes use of
reference annotations containing the time position of target sounds that
along with the corresponding sound files, serves as a training data set
in which to evaluate the performance of detection routines. This can be
done with routine outputs imported from other software as well as
detection routines run within the \emph{ohun} package itself. The
package also provides a set of functions to explore acoustic data sets
and organize them in a format amenable for detection analyses. In
addition, it offers implementations of two automatic detection methods
commonly used in bioacoustic analysis: energy-based detection and
template-based detection
(\protect\hyperlink{ref-mellinger2000}{Mellinger \& Clark 2000};
\protect\hyperlink{ref-Charif2010}{Charif \emph{et al.} 2010};
\protect\hyperlink{ref-aide2013}{Aide \emph{et al.} 2013};
\protect\hyperlink{ref-avi2014}{Bioacoustics 2014};
\protect\hyperlink{ref-Hafner2015}{Hafner \& Katz 2015}). We explain how
to explore and format acoustic data sets, how sound event detection
routines can be evaluated and then showcase the package usage with study
cases on male Zebra-finch songs (\emph{Taenopygia gutata}) and Spix's
disc-winged bat calls (\emph{Thyroptera tricolor}) which represent
different recording settings (lab and flight cages) and signal types
(sonic mating sounds and ultrasonic social calls).

\hypertarget{formatting-acoustic-data-sets}{%
\subsection{Formatting acoustic data
sets}\label{formatting-acoustic-data-sets}}

The format and size of the acoustic data to be analyzed is important to
validate and standardize in order to avoid downstream errors, as well as
to inform expectations for computational time performance. Several
functions in \emph{ohun} can facilitate double-checking the format of
acoustic data sets prior to automatic detection. The function
\texttt{feature\_acoustic\_data} prints a summary of the duration, size
and format of all the recordings in a folder. Here we explore the
acoustic data set of zebra finch's songs (Supporting Information):

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# working directory}
\NormalTok{path\_zebra\_finch }\OtherTok{\textless{}{-}} \StringTok{"path\_to\_zebra\_finch\_files"}

\FunctionTok{feature\_acoustic\_data}\NormalTok{(}\AttributeTok{path =}\NormalTok{ path\_zebra\_finch)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Features of the acoustic data set in '/home/m/Dropbox/Projects/ohun_paper/data/raw/taeniopygia':
\end{verbatim}

\begin{verbatim}
## 
## * 18 sound files
\end{verbatim}

\begin{verbatim}
## 
## * 1 file format(s) (.wav (18))
\end{verbatim}

\begin{verbatim}
## 
## * 1 sampling rate(s) (44.1 kHz (18))
\end{verbatim}

\begin{verbatim}
## 
## * 1 bit depth(s) (16 bits (18))
\end{verbatim}

\begin{verbatim}
## 
## * 1 number of channels (1 channel(s) (18))
\end{verbatim}

\begin{verbatim}
## 
## * File duration range: 2.63-21.76 s (mean: 12.22 s)
\end{verbatim}

\begin{verbatim}
## 
## * File size range: 0.23-1.92 MB (mean: 1.08 MB)
\end{verbatim}

\begin{verbatim}
## 
##  (detailed information by sound file can be obtained with 'warbleR::info_sound_files()')
\end{verbatim}

In this case all recordings have the same format (.wav files, 44.1 kHz
sampling rate, 16 bit resolution and a single channel). We can also
check the duration and size of files. Format information is important as
some tuning parameters of detection routines can behave differently
depending on file format (\emph{e.g.} time window size can be affected
by sampling rate) or simply because some software might only work on
specific sound file formats. In addition, long sound files could be hard
to analyze in some computers and might have to be split into shorter
clips. In the latter case the function \texttt{split\_acoustic\_data}
can be used to produce those clips:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{split\_info }\OtherTok{\textless{}{-}} \FunctionTok{split\_acoustic\_data}\NormalTok{(}\AttributeTok{path =}\NormalTok{ path\_zebra\_finch, }\AttributeTok{sgmt.dur =} \DecValTok{5}\NormalTok{)}

\FunctionTok{head}\NormalTok{(split\_info)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c}
\hline
original.sound.files & sound.files & start & end\\
\hline
Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & Ag13\_43421.27975590\_11\_17\_7\_46\_15-1.wav & 0 & 5.000\\
\hline
Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & Ag13\_43421.27975590\_11\_17\_7\_46\_15-2.wav & 5 & 10.000\\
\hline
Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & Ag13\_43421.27975590\_11\_17\_7\_46\_15-3.wav & 10 & 15.000\\
\hline
Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & Ag13\_43421.27975590\_11\_17\_7\_46\_15-4.wav & 15 & 20.000\\
\hline
Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & Ag13\_43421.27975590\_11\_17\_7\_46\_15-5.wav & 20 & 21.761\\
\hline
Blk109Brn\_43559.32349131\_4\_4\_8\_59\_9.wav & Blk109Brn\_43559.32349131\_4\_4\_8\_59\_9-1.wav & 0 & 5.000\\
\hline
\end{tabular}
\end{table}

The output shows the time segments in the original sound files to which
the clips belong to. If an annotation table is supplied (argument `X')
the function will adjust the annotations so they refer to the position
of the sounds in the clips. This can be helpful when reference tables
have been annotated on the original sound files.

Annotations can also be explored using the function
\texttt{feature\_reference}, which returns the mean and range of signal
duration and gap duration (time intervals between selections), bottom
and top frequency and of the number of annotations by sound file. If the
path to the sound files is supplied, then the duty cycle (the fraction
of a sound file corresponding to sounds) and peak amplitude (the highest
amplitude in a detection) are also returned:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# read reference annotations}
\NormalTok{manual\_ref\_tae }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\FunctionTok{file.path}\NormalTok{(path\_zebra\_finch, }\StringTok{"manual\_selections\_Taeniopygia.csv"}\NormalTok{))}

\CommentTok{\# explore annotations}
\FunctionTok{feature\_reference}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ manual\_ref\_tae, }\AttributeTok{path =}\NormalTok{ path\_zebra\_finch)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                  min   mean     max
## sel.duration   15.54 103.30  319.41
## gap.duration   80.19 352.26 3024.23
## annotations     2.00  32.83   66.00
## duty.cycle      0.09   0.29    0.61
## peak.amplitude 43.28  65.19   88.69
## bottom.freq     0.50   0.50    0.50
## top.freq       10.00  10.00   10.00
\end{verbatim}

\hypertarget{diagnosing-detection-performance}{%
\subsection{Diagnosing detection
performance}\label{diagnosing-detection-performance}}

The \emph{ohun} package makes use of signal detection theory indices to
evaluate detection performance. Signal detection theory deals with the
process of recovering sounds (\emph{i.e.} target sounds) from background
noise (not necessarily acoustic noise) and it's widely used for
optimizing this decision making process in the presence of uncertainty
(\protect\hyperlink{ref-hossin2015}{Hossin \& Sulaiman 2015}). During a
detection routine, the detected `items' can be classified into 4
classes: true positives (TPs, target sounds correctly identified as
signal), false positives (FPs, noise incorrectly identified as
`signal'), false negatives (FNs, sounds incorrectly identified as noise)
and true negatives (TNs, background noise correctly identified as
noise). However, TNs cannot always be easily defined in the context of
sound event detection as noise might not be partitioned in discrete
units. Hence, the package makes use of TPs, FPs and FNs to calculate
three additional indices that can further assist with evaluating the
performance of a detection routine and are widely used in the context of
sound event detection (\protect\hyperlink{ref-knight2017}{Knight
\emph{et al.} 2017}): recall (correct detections relative to total
detections), precision (proportion of target sounds that were correctly
detected) and F1 score (combined recall and precision as the harmonic
mean of these two, providing a single value for evaluating performance,
a.k.a. F-score, F-measure or Dice similarity coefficient).

A perfect detection will have no false positives or false negatives,
which will result in both recall and precision equal to 1. However,
perfect detection cannot always be achieved. Therefore, some compromise
between detecting most target sounds plus some noise and excluding noise
but missing target sounds might be warranted. These indices provide a
useful framework for diagnosing and optimizing the performance of a
detection routine. Researchers can identify an appropriate balance
between these two extremes will be given by the relative costs of
missing sounds and mistaking noise for target sounds in the context of
their specific study goals.

\emph{ohun} offers tools to evaluate the performance of sound event
detection methods based on the indices described above. To accomplish
this, annotations derived from a detection routine are compared against
a reference table containing the time position of all target sounds in
the sound files. For instance, the following code evaluates a routine
run in Raven Pro 1.6 (\protect\hyperlink{ref-Charif2010}{Charif \emph{et
al.} 2010}) using the ``band limited energy detector'' option (minimum
frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.03968
s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a
subset of the zebra finch recordings described below:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{raven\_detec }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{"combined\_raven\_detection.csv"}\NormalTok{)}

\FunctionTok{diagnose\_detection}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ manual\_ref\_tae, }\AttributeTok{detection =}\NormalTok{ raven\_detec, }\AttributeTok{by =} \StringTok{"tuning\_parameters"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{raven\_detec }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{"./data/processed/combined\_raven\_detection.csv"}\NormalTok{)}

\FunctionTok{diagnose\_detection}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ manual\_ref\_tae, }\AttributeTok{detection =}\NormalTok{ raven\_detec, }\AttributeTok{by =} \StringTok{"tuning\_parameters"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção
\end{verbatim}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c|c|c}
\hline
tuning\_parameters & total.detections & true.positives & false.positives & false.negatives & split.positives & merged.positives & overlap.to.true.positives & recall & precision & f1.score\\
\hline
band: 1-10 kHz; sep: 0.005 s & 489 & 489 & 85 & 107 & 33 & 99 & 0.911 & 0.820 & 0.852 & 0.836\\
\hline
band: 1-10 kHz; sep: 0.02322 s & 79 & 79 & 76 & 512 & 0 & 556 & 0.994 & 0.134 & 0.510 & 0.212\\
\hline
band: 1-15 kHz; sep: 0.02322 s & 430 & 430 & 202 & 161 & 21 & 236 & 0.927 & 0.728 & 0.680 & 0.703\\
\hline
band: 1-22 kHz; sep: 0.02322.txt s & 494 & 494 & 526 & 101 & 51 & 153 & 0.870 & 0.830 & 0.484 & 0.612\\
\hline
\end{tabular}
\end{table}

The output shows the indices described above. Plus three additional
metrics specific to sound event detection: `split positives', `merge
positives' and `overlap to true positives'. `Split positives' is the
number of reference sounds overlapped by more than one detection, `merge
positives' the number of detected sounds overlapping with another
detection and `overlap to true positives' quantifies the mean overlap
between detections and reference signal (1 means complete overlap). The
function also allows to detail those indices by sound file. Here we show
the first 10 files:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{diag\_raven }\OtherTok{\textless{}{-}} \FunctionTok{diagnose\_detection}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ manual\_ref\_tae, }\AttributeTok{detection =}\NormalTok{ raven\_detec, }\AttributeTok{by =} \StringTok{"tuning\_parameters"}\NormalTok{, }\AttributeTok{by.sound.file =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{head}\NormalTok{(diag\_raven, }\DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c|c|c|c}
\hline
tuning\_parameters & sound.files & total.detections & true.positives & false.positives & false.negatives & split.positives & merged.positives & overlap.to.true.positives & recall & precision & f1.score\\
\hline
band: 1-10 kHz; sep: 0.005 s & Ag13\_43421.27975590\_11\_17\_7\_46\_15.wav & 40 & 35 & 5 & 0 & 0 & 0 & 0.988 & 1.000 & 0.875 & 0.933\\
\hline
band: 1-10 kHz; sep: 0.005 s & BRN7\_43435.27985312\_12\_1\_7\_46\_25.wav & 34 & 27 & 7 & 24 & 0 & 21 & 0.980 & 0.529 & 0.794 & 0.635\\
\hline
band: 1-10 kHz; sep: 0.005 s & Blk109Brn\_43559.32349131\_4\_4\_8\_59\_9.wav & 30 & 29 & 1 & 5 & 0 & 6 & 0.920 & 0.853 & 0.967 & 0.906\\
\hline
band: 1-10 kHz; sep: 0.005 s & DB118\_43568.33139566\_4\_13\_9\_12\_19.wav & 18 & 18 & 0 & 1 & 0 & 2 & 0.989 & 0.947 & 1.000 & 0.973\\
\hline
band: 1-10 kHz; sep: 0.005 s & DB15HP\_43450.29217192\_12\_16\_8\_6\_57.wav & 23 & 21 & 2 & 0 & 0 & 0 & 0.989 & 1.000 & 0.913 & 0.955\\
\hline
band: 1-10 kHz; sep: 0.005 s & DB7\_43357.58119484\_9\_14\_16\_8\_39.wav & 18 & 18 & 0 & 1 & 0 & 2 & 0.941 & 0.947 & 1.000 & 0.973\\
\hline
band: 1-10 kHz; sep: 0.005 s & DG124DB\_43559.30160960\_4\_4\_8\_22\_40.wav & 39 & 29 & 10 & 0 & 0 & 0 & 0.986 & 1.000 & 0.744 & 0.853\\
\hline
band: 1-10 kHz; sep: 0.005 s & GRY37HP\_43442.27431670\_12\_8\_7\_37\_11.wav & 54 & 48 & 6 & 6 & 0 & 8 & 0.936 & 0.889 & 0.889 & 0.889\\
\hline
band: 1-10 kHz; sep: 0.005 s & Gold183\_43555.5549813\_3\_31\_1\_32\_29.wav & 2 & 2 & 0 & 0 & 0 & 0 & 1.000 & 1.000 & 1.000 & 1.000\\
\hline
band: 1-10 kHz; sep: 0.005 s & Gry35HP\_43455.29800260\_12\_21\_8\_16\_40.wav & 18 & 18 & 0 & 8 & 0 & 13 & 0.995 & 0.692 & 1.000 & 0.818\\
\hline
\end{tabular}
\end{table}

Diagnostics from routines using different tuning parameters can be used
to identify the parameter values that optimize detection. These process
of evaluating different routines for detection optimization is
incorporated into the two signal detection approaches provided natively
by \emph{ohun}, which we depict in the following section. Note that the
detection with Raven Pro does not necessarily reflect the best
performance of this software and has been included only as an example on
evaluating detection from external sources, rather than a direct
comparison of performance between Raven Pro and \emph{ohun}.

\hypertarget{signal-detection-with-ohun}{%
\subsection{\texorpdfstring{Signal detection with
\emph{ohun}}{Signal detection with ohun}}\label{signal-detection-with-ohun}}

The package offers two methods for automated signal detection:
template-based and energy-based detection. These methods are better
suited for highly stereotyped or good signal-to-noise ratio (SNR)
sounds, respectively. If the target sounds do not fit these
requirements, more elaborate methods (i.e.~machine/deep learning
approaches) are warranted.

\hypertarget{study-cases}{%
\subsection{Study cases}\label{study-cases}}

\hypertarget{template-detection-on-ultrasonic-social-calls-of-spixs-disc-winged-bats}{%
\subsubsection{Template detection on ultrasonic social calls of Spix's
disc-winged
bats}\label{template-detection-on-ultrasonic-social-calls-of-spixs-disc-winged-bats}}

We recorded 30 individuals of Spix's disc-winged bats (\emph{Thyroptera
tricolor}) at Baru Biological Station, in south-western Costa Rica in
January 2020. Bats were captured at their roosting sites (furled leaves
of Zingeberaceae plants). Each individual bat was released in a large
flight cage (9 x 4 x 3 m) for a 5-minute period and their ultrasonic
inquiry calls were recorded using a condenser microphone (CM16, Avisoft
Bioacoustics, Glienike/Nordbahn, Germany) through an Avisoft
UltraSoundGate 116Hm plugged into a laptop computer running
Avisoft-Recorder software. Recordings were made at a sampling rate of
500 kHz and an amplitude resolution of 16 bits.

Recordings were manually annotated using Raven Pro 1.6
(\protect\hyperlink{ref-Charif2010}{Charif \emph{et al.} 2010}).
Annotations were created by visual inspection of spectrograms, in which
the start and end of sounds were determined by the location of the
continuous traces of power spectral entropy of the target sounds. A
total of 644 calls were annotated (\textasciitilde21 calls per
recording). Annotations were made with a time window of 200 samples and
70\% of overlap, and were then imported into R using the package Rraven
(\protect\hyperlink{ref-Araya-Salas2017e}{Araya-Salas 2020}).

Inquiry calls of Spix's disc-winged bats are structurally stereotyped
(\protect\hyperlink{ref-Chaverri2010b}{Chaverri \emph{et al.} 2010}).
Most variation is found among individuals, although the basic form of a
short, downward broadband frequency modulation is always shared (Fig.
BAT-SPECTRO, Araya-Salas et al 2021 ontogeny).

\begin{verbatim}
## Error in knitr::include_graphics(file.path(rootdir, "data/raw/thyroptera/Catalog_p1.jpeg")): Cannot find the file(s): "../data/raw/thyroptera/Catalog_p1.jpeg"
\end{verbatim}

Template-based detection is a useful approach when there are minimal
structural differences in the target sounds
(\protect\hyperlink{ref-knight2017}{Knight \emph{et al.} 2017};
e.g.~when signals are produced in a highly stereotyped manner,
\protect\hyperlink{ref-balantic2020}{Balantic \& Donovan 2020}). It uses
spectrographic cross-correlation to find sounds resembling an example
target sounds (i.e.~template) across sound files, applying a correlation
threshold to separate detections from background noise. We used this
approach in \emph{ohun} to detect inquiry calls. To do this, we tested
the performance of three acoustic templates on a training subset of five
sound files. First we used the function \texttt{get\_templates} to find
several sounds representative of the variation in signal structure. This
function measures several spectral parameters which are then summarized
using Principal Component Analysis. The first two components are used to
project the acoustic space. In this space the function defines
sub-spaces as equal-size slices of a sphere centered at the centroid of
the acoustic space. Templates are then selected as those closer to the
centroid within each of the sub-spaces, including the centroid for the
entire acoustic space. The user needs to define the number of sub-spaces
in which the acoustic space will be split.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# read manual annotations}
\NormalTok{manual\_ref\_thy }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{"manual\_annotations\_thyroptera.csv"}\NormalTok{)}

\CommentTok{\# get random subset of 5 sound files for training}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\NormalTok{train\_files }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(}\FunctionTok{unique}\NormalTok{(manual\_ref\_thy}\SpecialCharTok{$}\NormalTok{sound.files), }\AttributeTok{size =} \DecValTok{5}\NormalTok{)}
\NormalTok{train\_ref }\OtherTok{\textless{}{-}}\NormalTok{ manual\_ref\_thy[manual\_ref\_thy}\SpecialCharTok{$}\NormalTok{sound.files }\SpecialCharTok{\%in\%}\NormalTok{ train\_files, ]}

\CommentTok{\# the rest for testin g}
\NormalTok{test\_files }\OtherTok{\textless{}{-}} \FunctionTok{setdiff}\NormalTok{(manual\_ref\_thy}\SpecialCharTok{$}\NormalTok{sound.files, train\_files)}
\NormalTok{test\_ref }\OtherTok{\textless{}{-}}\NormalTok{ manual\_ref\_thy[manual\_ref\_thy}\SpecialCharTok{$}\NormalTok{sound.files }\SpecialCharTok{\%in\%}\NormalTok{ test\_files, ]}

\CommentTok{\# find templates}
\NormalTok{templates }\OtherTok{\textless{}{-}} \FunctionTok{get\_templates}\NormalTok{(train\_ref, }\AttributeTok{path =}\NormalTok{ data\_path,  }\AttributeTok{bp =} \FunctionTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{50}\NormalTok{), }\AttributeTok{ovlp =} \DecValTok{70}\NormalTok{, }\AttributeTok{wl =} \DecValTok{200}\NormalTok{, }\AttributeTok{n.sub.spaces =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## The first 2 principal components explained 0.5 of the variance
\end{verbatim}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/unnamed-chunk-12-1.pdf}
\caption{FIG. ACOUSTIC-SPACE. Acoustic space defined as the the first
two components of a Principal Component Analysis on spectrographic
parameters. Templates are selected as those closer to the centroid for
each of the sub-spaces. Gray dashed lines delimit the region of
sub-spaces. Yellow circles around points highlight the position of the
signals selected as templates.}
\end{figure}

The output of the \texttt{get\_templates} function includes an acoustic
space plot (FIG. ACOUSTIC-SPACE) in which the position of the sounds
selected as templates is highlighted. Users can also provide their own
acoustic space dimensions (argument `acoustic.space'). In the following
code we used those templates for detecting bat social calls. The code
iterates a template-based detection on the training data set across a
range of correlation thresholds, in order to find the combination of
threshold and template with the best performance.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# get correlation vectors}
\NormalTok{corr\_templ\_train }\OtherTok{\textless{}{-}} \FunctionTok{template\_correlator}\NormalTok{(}
  \AttributeTok{templates =}\NormalTok{ templates,}
    \AttributeTok{path =}\NormalTok{ data\_path, }
  \AttributeTok{files =} \FunctionTok{unique}\NormalTok{(train\_ref}\SpecialCharTok{$}\NormalTok{sound.files), }
  \AttributeTok{hop.size =} \DecValTok{10}\NormalTok{, }
  \AttributeTok{ovlp =} \DecValTok{70}
\NormalTok{  )}

\CommentTok{\# evaluate detection for different correlation thresholds}
\NormalTok{opt\_detec\_train }\OtherTok{\textless{}{-}} \FunctionTok{optimize\_template\_detector}\NormalTok{(}
  \AttributeTok{reference =}\NormalTok{ train\_ref, }
  \AttributeTok{template.correlations =}\NormalTok{ corr\_templ\_train,}
  \AttributeTok{threshold =} \FunctionTok{seq}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.01}\NormalTok{)}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

Note that the correlation vectors are estimated first (i.e.~vectors of
correlation values across sound files, \texttt{template\_correlator()})
and then the correlation thresholds are optimized on these vectors
(\texttt{optimize\_template\_detector()}). The output of
\texttt{optimize\_template\_detector()} contains the detection
performance indices for each combination of templates and thresholds.
Table TEMPLATE PERFORMANCE shows the two highest performance runs for
each template.

\begin{table}

\caption{\label{tab:unnamed-chunk-15}TABLE TEMPLATE PERFORMANCE. Performance diagnostic of template-based detections using four templates across several threshold values. Only the two highest performance iterations for each template are shown.}
\centering
\begin{tabular}[t]{r|l|r|r|r|r|r|r}
\hline
threshold & templates & true.positives & false.positives & false.negatives & recall & precision & f1.score\\
\hline
0.45 & centroid & 81 & 1 & 0 & 1.000 & 0.988 & 0.994\\
\hline
0.50 & centroid & 80 & 0 & 1 & 0.988 & 1.000 & 0.994\\
\hline
0.50 & templ-1 & 77 & 1 & 4 & 0.951 & 0.987 & 0.969\\
\hline
0.45 & templ-1 & 80 & 5 & 1 & 0.988 & 0.941 & 0.964\\
\hline
0.40 & templ-2 & 79 & 0 & 2 & 0.975 & 1.000 & 0.988\\
\hline
0.35 & templ-2 & 80 & 2 & 1 & 0.988 & 0.976 & 0.982\\
\hline
0.45 & templ-3 & 77 & 2 & 4 & 0.951 & 0.975 & 0.963\\
\hline
0.40 & templ-3 & 79 & 5 & 2 & 0.975 & 0.940 & 0.958\\
\hline
\end{tabular}
\end{table}

We can explore the performance of each template in more detail by
looking at the change in F1 score across thresholds (FIG. THRESHOLD vs
F1.SCORE).

\begin{figure}

{\centering \includegraphics{manuscript_files/figure-latex/unnamed-chunk-16-1}

}

\caption{FIG. THRESHOLD vs F1.SCORE. Shows the changes in F1 score across the range of threshold values}\label{fig:unnamed-chunk-16}
\end{figure}

In this example the ``centroid'' template, produced the best performance
(TABLE TEMPLATE PERFORMANCE; FIG. THRESHOLD vs F1.SCORE). Hence, we will
use this template for detecting calls on the rest of the data. The
following code extracts this template from the reference annotation
table and use it to find inquiry calls on the testing data set:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# get correlation vectors for test files}
\NormalTok{corr\_templ\_test }\OtherTok{\textless{}{-}} \FunctionTok{template\_correlator}\NormalTok{(}
  \AttributeTok{templates =}\NormalTok{ templates[templates}\SpecialCharTok{$}\NormalTok{sound.file }\SpecialCharTok{==} \StringTok{"centroid"}\NormalTok{, ],}
    \AttributeTok{path =}\NormalTok{ data\_path, }\AttributeTok{files =} \FunctionTok{unique}\NormalTok{(test\_ref}\SpecialCharTok{$}\NormalTok{sound.files), }
  \AttributeTok{hop.size =} \DecValTok{10}\NormalTok{, }
  \AttributeTok{ovlp =} \DecValTok{70}
\NormalTok{  )}

\CommentTok{\# detect on test files}
\NormalTok{detec\_test }\OtherTok{\textless{}{-}} \FunctionTok{template\_detector}\NormalTok{(}
  \AttributeTok{template.correlations =}\NormalTok{ corr\_templ\_test,}
    \AttributeTok{threshold =} \FloatTok{0.45}
\NormalTok{  )}

\FunctionTok{diagnose\_detection}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ test\_ref, }\AttributeTok{detection =}\NormalTok{ detec\_test)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção
\end{verbatim}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c|c}
\hline
total.detections & true.positives & false.positives & false.negatives & split.positives & merged.positives & overlap.to.true.positives & recall & precision & f1.score\\
\hline
522 & 522 & 10 & 41 & 2 & 32 & 0.907 & 0.927 & 0.981 & 0.953\\
\hline
\end{tabular}
\end{table}

The last line of codes evaluates the detection on the test data set,
which shows a good performance for both recall and precision.

\hypertarget{energy-based-detection-on-zebra-finch-vocalizations}{%
\subsubsection{Energy-based detection on zebra finch
vocalizations}\label{energy-based-detection-on-zebra-finch-vocalizations}}

We used recordings from 18 zebra finch males recorded at obtained from
the Rockefeller University Field Research Center Song Library
(\url{http://ofer.sci.ccny.cuny.edu/songs},
\protect\hyperlink{ref-Tchernichovski2021}{Tchernichovski \emph{et al.}
2021}). Recordings contain undirected vocalizations (songs or calls) of
single males recorded in sound attenuation chambers using Sound Analysis
Pro. Zebra finch vocalizations are composed of multiple elements
(i.e.~distinct patterns of continuous traces of power spectral entropy
in the spectrogram separated by time gaps) that can vary importantly in
key features as duration and frequency range (ZEBRAFINCH-SPECTRO) and
are clearly not as stereotyped as the Spix's disc-winged bats. However,
as the recorded sounds show a good signal-to-noise ratio, such that
signals in each recording can potentially be detected using an
energy-based approach that does not rely on matching the acoustic
structure of a template.

\begin{verbatim}
## Error in knitr::include_graphics(file.path(rootdir, "data/raw/taeniopygia/same.time-Catalog_p1.jpeg")): Cannot find the file(s): "../data/raw/taeniopygia/same.time-Catalog_p1.jpeg"
\end{verbatim}

Reference annotations were made manually on the oscillogram with the
spectrogram and audio as a guide using Raven Lite 2.0.1 (Cornell Lab of
Ornithology). The following code loads the reference annotations and
split them into two data sets for training (3 sound files) and testing
(15 sound files):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{manual\_ref\_tae }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{"manual\_selections\_Taeniopygia.csv"}\NormalTok{)}

\FunctionTok{set.seed}\NormalTok{(}\DecValTok{450}\NormalTok{)    }
\NormalTok{train\_files }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(}\FunctionTok{unique}\NormalTok{(manual\_ref\_tae}\SpecialCharTok{$}\NormalTok{sound.files), }\DecValTok{3}\NormalTok{)    }
\NormalTok{test\_files }\OtherTok{\textless{}{-}} \FunctionTok{setdiff}\NormalTok{(manual\_ref\_tae}\SpecialCharTok{$}\NormalTok{sound.files, train\_files)}

\NormalTok{train\_ref }\OtherTok{\textless{}{-}}\NormalTok{ manual\_ref\_tae[manual\_ref\_tae}\SpecialCharTok{$}\NormalTok{sound.files }\SpecialCharTok{\%in\%}\NormalTok{ train\_files, ]}
\NormalTok{test\_ref }\OtherTok{\textless{}{-}}\NormalTok{ manual\_ref\_tae[manual\_ref\_tae}\SpecialCharTok{$}\NormalTok{sound.files }\SpecialCharTok{\%in\%}\NormalTok{ test\_files, ]}
\end{Highlighting}
\end{Shaded}

The detection parameters can be optimized using the function
\texttt{optimize\_energy\_detector}. This function runs a detection for
all possible combinations of tuning parameters. The code below tries
three minimum duration and maximum duration values and two hold time
values (which merges sounds within the specified time interval):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{opt\_det\_train }\OtherTok{\textless{}{-}} \FunctionTok{optimize\_energy\_detector}\NormalTok{(}
  \AttributeTok{reference =}\NormalTok{ train\_ref, }
  \AttributeTok{files =}\NormalTok{ train\_files, }
  \AttributeTok{threshold =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{5}\NormalTok{), }
  \AttributeTok{hop.size =} \FloatTok{11.6}\NormalTok{, }
  \AttributeTok{smooth =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{10}\NormalTok{), }
  \AttributeTok{hold.time =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{5}\NormalTok{), }
  \AttributeTok{min.duration =} \FunctionTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{25}\NormalTok{), }
  \AttributeTok{max.duration =} \FunctionTok{c}\NormalTok{(}\DecValTok{275}\NormalTok{, }\DecValTok{300}\NormalTok{, }\DecValTok{325}\NormalTok{), }
  \AttributeTok{bp =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\DecValTok{10}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The output (\texttt{opt\_det\_train}) shows the performance indices for
each of those combinations. Hhere we show the 10 combinations with the
highest F1 score:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# subset with highest performance}
\NormalTok{opt\_det\_train }\OtherTok{\textless{}{-}}\NormalTok{ opt\_det\_train[}\FunctionTok{order}\NormalTok{(opt\_det\_train}\SpecialCharTok{$}\NormalTok{f1.score, }\AttributeTok{decreasing =} \ConstantTok{TRUE}\NormalTok{), ]}

\FunctionTok{head}\NormalTok{(opt\_det\_train, }\DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c|c|c}
\hline
threshold & smooth & hold.time & min.duration & max.duration & true.positives & false.positives & false.negatives & recall & precision & f1.score\\
\hline
1 & 5 & 5 & 25 & 300 & 105 & 9 & 0 & 1.000 & 0.921 & 0.959\\
\hline
1 & 5 & 5 & 25 & 325 & 105 & 9 & 0 & 1.000 & 0.921 & 0.959\\
\hline
1 & 5 & 0 & 25 & 300 & 105 & 12 & 0 & 1.000 & 0.897 & 0.946\\
\hline
1 & 5 & 0 & 25 & 325 & 105 & 12 & 0 & 1.000 & 0.897 & 0.946\\
\hline
1 & 10 & 0 & 25 & 300 & 105 & 13 & 0 & 1.000 & 0.890 & 0.942\\
\hline
1 & 10 & 0 & 25 & 325 & 105 & 13 & 0 & 1.000 & 0.890 & 0.942\\
\hline
1 & 5 & 5 & 15 & 300 & 105 & 15 & 0 & 1.000 & 0.875 & 0.933\\
\hline
1 & 5 & 5 & 15 & 325 & 105 & 15 & 0 & 1.000 & 0.875 & 0.933\\
\hline
5 & 5 & 5 & 15 & 275 & 96 & 8 & 9 & 0.914 & 0.923 & 0.919\\
\hline
5 & 5 & 5 & 15 & 300 & 96 & 8 & 9 & 0.914 & 0.923 & 0.919\\
\hline
\end{tabular}
\end{table}

We now can use the tuning parameter values that yielded the best
performance to detect sounds on the test data set:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{best\_param }\OtherTok{\textless{}{-}}\NormalTok{ opt\_det\_train[}\FunctionTok{which.max}\NormalTok{(opt\_det\_train}\SpecialCharTok{$}\NormalTok{f1.score), ]}

\NormalTok{det\_test }\OtherTok{\textless{}{-}} \FunctionTok{energy\_detector}\NormalTok{(}
  \AttributeTok{files =}\NormalTok{ test\_files, }
  \AttributeTok{threshold =}\NormalTok{ best\_param}\SpecialCharTok{$}\NormalTok{threshold, }
  \AttributeTok{hop.size =} \FloatTok{11.6}\NormalTok{, }
  \AttributeTok{smooth =}\NormalTok{ best\_param}\SpecialCharTok{$}\NormalTok{smooth, }
  \AttributeTok{hold.time =}\NormalTok{ best\_param}\SpecialCharTok{$}\NormalTok{hold.time, }
  \AttributeTok{min.duration =}\NormalTok{ best\_param}\SpecialCharTok{$}\NormalTok{min.duration, }
  \AttributeTok{max.duration =}\NormalTok{ best\_param}\SpecialCharTok{$}\NormalTok{max.duration, }
  \AttributeTok{bp =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\DecValTok{10}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

As our reference annotations include all sound files we can evaluate the
performance of the detection on the test set as well:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{diagnose\_detection}\NormalTok{(}\AttributeTok{reference =}\NormalTok{ test\_ref, }\AttributeTok{detection =}\NormalTok{ det\_test, }\AttributeTok{by.sound.file =} \ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção

## Warning in FUN(X[[i]], ...): NAs introduzidos por coerção
\end{verbatim}

\begin{table}
\centering
\begin{tabular}[t]{c|c|c|c|c|c|c|c|c|c}
\hline
total.detections & true.positives & false.positives & false.negatives & split.positives & merged.positives & overlap.to.true.positives & recall & precision & f1.score\\
\hline
463 & 463 & 27 & 27 & 4 & 32 & 0.97 & 0.945 & 0.945 & 0.945\\
\hline
\end{tabular}
\end{table}

~

The performance on the test data set was also acceptable with a F1 score
of 0.95. Note that in the example we used a small subset of sound files
for training. More training data might be needed for optimizing a
detection routine on larger data sets or recordings with more variable
sounds or background noise levels.

\hypertarget{additional-tools}{%
\subsubsection{Additional tools}\label{additional-tools}}

The \emph{ohun} package offers additional tools to simplify sound event
detection. Detected sounds can be labeled as false or true positives
with the function `label\_detection'. This allows users to explore the
structure of false positives and figure out ways to exclude them. The
function `filter\_detections' can be used to remove ambiguous sounds
(i.e.~those labeled as split or merged detections) keeping only those
that maximize a specific criterion (i.e.~the highest template
correlation). Finally, note that several templates representing the
range of variation in signal structure can be used to detect
semi-stereotyped sounds when running template-based detection
(`template\_detection' function).

\hypertarget{discusion}{%
\subsection{Discusion}\label{discusion}}

Here we have shown how to evaluate the performance of sound event
detection routines using the package \emph{ohun}. The package can
evaluate detection outputs imported from other software, as well as its
own detection routines. The latter can be iterated over combinations of
tuning parameters in order to find those values that optimize detection.
Despite signal detection indices being commonly reported when presenting
new automatic detection methods, to our knowledge widely applicable
performance evaluating routines have not been made available in a free,
open source platform. Providing a common framework for the evaluation of
sound event detection can simplify comparing the performance of
different tools and the selection of those tools better suited to a
given research question and study system. The tools offered by
\emph{ohun} for diagnosing detection performance should not be
necessarily limited to acoustic data. \emph{ohun} can also be used for
cases in which the time of occurrence of discrete events needs to be
identified such as detecting specific behaviors in video analysis of
animal motor activity (\emph{e.g.}
\protect\hyperlink{ref-sturman2020}{Sturman \emph{et al.} 2020};
\protect\hyperlink{ref-hsu2021}{Hsu \& Yttri 2021};
\protect\hyperlink{ref-bohnslav2021}{Bohnslav \emph{et al.} 2021}). The
detection of such motor events in video can also be evaluated and
optimized in comparison to a reference annotation, as we have shown
above for acoustic signals.

The \emph{ohun} package provides two native detection methods:
template-based and energy-based detection. Compared to new deep learning
approaches for finding the occurrence of sound events, the two native
methods are relatively simple tools. However, these methods have been
widely used by the bioacoustic community
(\protect\hyperlink{ref-mellinger2000}{Mellinger \& Clark 2000};
\protect\hyperlink{ref-Charif2010}{Charif \emph{et al.} 2010};
\protect\hyperlink{ref-aide2013}{Aide \emph{et al.} 2013};
\protect\hyperlink{ref-avi2014}{Bioacoustics 2014};
\protect\hyperlink{ref-Hafner2015}{Hafner \& Katz 2015}) and can reach
adequate performance under the appropriate conditions, as we have shown
in our two study cases and has been shown elsewhere
(\protect\hyperlink{ref-knight2017}{Knight \emph{et al.} 2017}). Deep
learning methods tend to require greater computational power, larger
training data sets and, in some cases, more complex training routines.
This might bring unnecessary difficulties when dealing with when less
challenging detection tasks. Therefore the availability of a wide range
of approaches can simplify finding the most appropriate tool for the
intricacies of a study system and research goals, as well as having
tools accessible for a broader community of researchers. The tools
offered in \emph{ohun} can also be used in a subsequent pipeline in
which detected sounds are further classified using more elaborated
discrimination algorithms (\protect\hyperlink{ref-balantic2020}{Balantic
\& Donovan 2020}). In addition, detection performance can be improved by
using acoustic structure measurements that can be used to distinguish
target from non-target sound events.

Detection routines can take a long time when working with large amounts
of acoustic data (e.g.~large recordings and/or many files). We provide
some additional tips that can help make a routine more time-efficient.
1. Always test procedures on small data subsets. Make sure that you
obtain decent results on a small subset of recordings before trying to
scale up the analysis. 2. Template-based detection is almost always
faster than energy-based detection. 3. Run routines in parallel.
Parallelization (i.e.~the ability to distribute tasks over several cores
in your computer) can significantly speed-up routines. All functions for
automatic detection and performance evaluation allow users to run
analysis in parallel (see \texttt{parallel} argument in those
functions). Hence a computer with several cores can be helpful for
improving efficiency. 4. Sampling rate matters. Detecting sounds on low
sampling rate files is faster, so we must avoid having Nyquist
frequencies much higher than the highest frequency of the target sounds.
5. Try using a computer with lots of RAM memory or a computer cluster
for working on large amounts of data. Lastly, we underscore that these
tips are not restricted to \emph{ohun} and can also be helpful to speed
up routines in other software packages.

There are some additional things to be considered when aiming to
automatically detect sound events. When running energy-based detection
routines try to use your knowledge about the signal structure to
determine the initial range of tunning parameters. This can be extremely
helpful for narrowing down possible parameter values. As a general rule,
if human observers have a hard time detecting where a target sound
occurs in a sound file, detection algorithms will likely yield low
detection performance. In cases in which occurrences are ambiguous low
performances are expected. Ensure reference annotations contain all
target sounds and only the target sounds, otherwise performance
optimization can be misleading as the performance of a given detection
method cannot be better than the reference itself. Lastly, avoid having
overlapping sounds or several sounds as a single detection (like a
multi-syllable vocalization) in the reference annotation when running an
energy-based detector, as they are likely to be identified as separated
units.

\hypertarget{acknowledgements}{%
\subsubsection{Acknowledgements}\label{acknowledgements}}

Nazareth Rojas, Silvia Chaves-Ramírez, Mariela Sánchez-Chavarría, Miriam
Gioiosa, Cristian Castillo-Salazar, and José Pablo Barrantes for their
help collecting acoustic data for Spix's disc-winged bats and Jorge
Elizondo for his help on manually annotating recordings. We also thank
the Centro Biológico Hacienda Barú for their continuous support of our
research and Mijail Rojas and Andrey Sequeira for support on early
stages of package development. This study was partly funded by a
CONARE-Max Planck grant from the Consejo Nacional de Rectores and the
research activity C0754 at Centro de Investigación en Neurociencias,
Universidad de Costa Rica. A.R-G. is supported by the Walt Halperin
Endowed Professorship and the Washington Research Foundation as
Distinguished Investigator.

\hypertarget{ethical-note}{%
\subsubsection{Ethical note}\label{ethical-note}}

All sampling protocols followed guidelines approved by the American
Society of Mammalogists for capture, handling and care of mammals
(\protect\hyperlink{ref-Sikes2016}{Sikes \emph{et al.} 2016}) and the
ASAB/ABS Guidelines for the use of animals in research. This study was
conducted in accordance with the ethical standards for animal welfare of
the Costa Rican Ministry of Environment and Energy, Sistema Nacional de
Áreas de Conservación, permit no. SINAC-ACOPAC-RES-INV-008-2017 (Decree
No.~32553-MINAE). Protocols were also approved by the University of
Costa Rica's Institutional Animal Care and Use Committee
(CICUA-42-2018).

\hypertarget{supporting-information}{%
\subsubsection{Supporting Information}\label{supporting-information}}

Supplementary information associated with this article is available in
XXXXXXXXXXXXXXXXX

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-aide2013}{}}%
Aide, T.M., Corrada-Bravo, C., Campos-Cerqueira, M., Milan, C., Vega, G.
\& Alvarez, R. (2013).
\href{https://doi.org/10.7717/PEERJ.103}{Real-time bioacoustics
monitoring and automated species identification}. \emph{PeerJ},
\textbf{2013}.

\leavevmode\vadjust pre{\hypertarget{ref-Araya-Salas2017e}{}}%
Araya-Salas, M. (2020). Rraven: Connecting r and raven bioacoustic
software. R package version 1.0.9.

\leavevmode\vadjust pre{\hypertarget{ref-Araya-Salas2017f}{}}%
Araya-Salas, M. \& Smith-Vidaurre, G. (2017). warbleR: An r package to
streamline analysis of animal acoustic signals. \emph{Methods in Ecology
and Evolution}, \textbf{8}, 184--191. Retrieved from
\url{http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12624/abstract}

\leavevmode\vadjust pre{\hypertarget{ref-balantic2020}{}}%
Balantic, C.M. \& Donovan, T.M. (2020). Statistical learning mitigation
of false positives from template-detected data in automated acoustic
wildlife monitoring. \emph{BIoacoustics}, \textbf{29}, 293--321.
Retrieved from
\url{https://www.tandfonline.com/action/journalInformation?journalCode=tbio20}

\leavevmode\vadjust pre{\hypertarget{ref-avi2014}{}}%
Bioacoustics, A. (2014). Avisoft SasLab pro.

\leavevmode\vadjust pre{\hypertarget{ref-Blumstein2011}{}}%
Blumstein, D.T., Mennill, D.J., Clemins, P., Girod, L., Yao, K.,
Patricelli, G., Deppe, J.L., Krakauer, A.H., Clark, C., Cortopassi,
K.A., Hanser, S.F., Mccowan, B., Ali, A.M. \& Kirschel, A.N.G. (2011).
\href{https://doi.org/10.1111/j.1365-2664.2011.01993.x}{Acoustic
monitoring in terrestrial environments using microphone arrays:
Applications, technological considerations and prospectus}.
\emph{Journal of Applied Ecology}, \textbf{48}, 758--767.

\leavevmode\vadjust pre{\hypertarget{ref-bohnslav2021}{}}%
Bohnslav, J.P., Wimalasena, N.K., Clausing, K.J., Dai, Y.Y.,
Yarmolinsky, D.A., Cruz, T., Kashlan, A.D., Chiappe, M.E., Orefice,
L.L., Woolf, C.J. \& Harvey, C.D. (2021).
\href{https://doi.org/10.7554/ELIFE.63377}{DeepEthogram, a machine
learning pipeline for supervised behavior classification from raw
pixels}. \emph{eLife}, \textbf{10}.

\leavevmode\vadjust pre{\hypertarget{ref-Burgdorf2011}{}}%
Burgdorf, J., Panksepp, J. \& Moskal, J.R. (2011).
\href{https://doi.org/10.1016/j.neubiorev.2010.11.011}{Frequency-modulated
50kHz ultrasonic vocalizations: A tool for uncovering the molecular
substrates of positive affect}. \emph{Neuroscience and Biobehavioral
Reviews}, \textbf{35}, 1831--1836.

\leavevmode\vadjust pre{\hypertarget{ref-Charif2010}{}}%
Charif, R., Waack, A. \& Strickman, L. (2010). \emph{Raven pro 1.4
user's manual.} Cornell Lab of Ornithology.

\leavevmode\vadjust pre{\hypertarget{ref-Chaverri2010b}{}}%
Chaverri, G., Gillam, E.H. \& Vonhof, M.J. (2010).
\href{https://doi.org/10.1098/rsbl.2009.0964}{Social calls used by a
leaf-roosting bat to signal location}. \emph{Biology Letters},
\textbf{6}, 441--444.

\leavevmode\vadjust pre{\hypertarget{ref-gibb2019}{}}%
Gibb, R., Browning, E., Glover-Kapfer, P., Jones, K.E. \& Jones, C.E.K.
(2019). Emerging opportunities and challenges for passive acoustics in
ecological assessment and monitoring. \emph{Wiley Online Library},
\textbf{10}, 169--185. Retrieved from
\url{https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/2041-210X.13101}

\leavevmode\vadjust pre{\hypertarget{ref-gew2019}{}}%
Gwee, C., Eaton, J., Garg, K. \& Alström, P. (2019). Cryptic diversity
in cyornis (aves: Muscicapidae) jungle-flycatchers flagged by simple
bioacoustic approaches. \emph{Zoological Journal}, \textbf{186}.
Retrieved from
\url{https://academic.oup.com/zoolinnean/article-abstract/186/3/725/5372980}

\leavevmode\vadjust pre{\hypertarget{ref-Hafner2015}{}}%
Hafner, S. \& Katz, J. (2015). monitoR: Acoustic template detection in
r. \emph{R package version 1.0.3}. Retrieved from
\href{http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htmURL:\%20http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htmURL:\%20http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htm}{http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htmURL:
http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htmURL:
http://www.uvm.edu/rsenr/vtcfwru/R/?Page=monitoR/monitoR.htm}

\leavevmode\vadjust pre{\hypertarget{ref-hossin2015}{}}%
Hossin, M. \& Sulaiman, M. (2015). A review on evaluation metrics for
data classification evaluations. \emph{International journal of data
mining}. Retrieved from
\url{https://www.academia.edu/download/37219940/5215ijdkp01.pdf}

\leavevmode\vadjust pre{\hypertarget{ref-hsu2021}{}}%
Hsu, A. \& Yttri, E. (2021). B-SOiD, an open-source unsupervised
algorithm for identification and fast prediction of behaviors.
\emph{Nature Communications}. Retrieved from
\url{https://www.nature.com/articles/s41467-021-25420-x}

\leavevmode\vadjust pre{\hypertarget{ref-knight2017}{}}%
Knight, E.C., Hannah, K.C., Foley, G.J., Scott, C.D., Brigham, R.M. \&
Bayne, E. (2017).
\href{https://doi.org/10.5751/ACE-01114-120214}{Recommendations for
acoustic recognizer performance assessment with application to five
common automated signal recognition programs}. \emph{Avian Conservation
and Ecology, Published online: Nov 27, 2017 \textbar{}
doi:10.5751/ACE-01114-120214}, \textbf{12}.

\leavevmode\vadjust pre{\hypertarget{ref-Kohler2017}{}}%
Köhler, J., Jansen, M., Rodríguez, A., Kok, P.J.R., Felipe, T.L.,
Emmrich, M., Glaw, F., Haddad, C.F.B., Mark-Oliver, R., Vences, M.,
Toledo, L.F., Emmrich, M., Glaw, F., Haddad, C.F.B., Rödel, M.O. \&
Vences, M. (2017).
\emph{\href{https://doi.org/10.11646/zootaxa.4251.1.1}{The use of
bioacoustics in anuran taxonomy: Theory, terminology, methods and
recommendations for best practice}}.

\leavevmode\vadjust pre{\hypertarget{ref-macmillan2004}{}}%
Macmillan, N.A. \& Creelman, C.D. (2004).
\href{https://doi.org/10.4324/9781410611147/DETECTION-THEORY-NEIL-MACMILLAN-DOUGLAS-CREELMAN}{Detection
theory: A user's guide: 2nd edition}. \emph{Detection Theory: A User's
Guide: 2nd edition}, 1--445.

\leavevmode\vadjust pre{\hypertarget{ref-Medina-Garcia2015a}{}}%
Medina-García, A., Araya-Salas, M. \& Wright, T.F. (2015). Does vocal
learning accelerate acoustic diversification? Evolution of contact calls
in neotropical parrots. \emph{Journal of Evolutionary Biology},
\textbf{28}, 1782--1792. Retrieved from
\url{http://doi.wiley.com/10.1111/jeb.12694}

\leavevmode\vadjust pre{\hypertarget{ref-mellinger2000}{}}%
Mellinger, D.K. \& Clark, C.W. (2000).
\href{https://doi.org/10.1121/1.429434}{Recognizing transient
low-frequency whale sounds by spectrogram correlation}. \emph{The
Journal of the Acoustical Society of America}, \textbf{107}, 3518--3529.

\leavevmode\vadjust pre{\hypertarget{ref-Odom2021}{}}%
Odom, K.J., Araya-Salas, M., Morano, J.L., Ligon, R.A., Leighton, G.M.,
Taff, C.C., Dalziell, A.H., Billings, A.C., Germain, R.R., Pardo, M.,
Andrade, L.G. de, Hedwig, D., Keen, S.C., Shiu, Y., Charif, R.A.,
Webster, M.S. \& Rice, A.N. (2021).
\href{https://doi.org/10.1111/brv.12695}{Comparative bioacoustics: A
roadmap for quantifying and comparing animal sounds across diverse
taxa}. \emph{Biological Reviews}, \textbf{96}, 1135--1159.

\leavevmode\vadjust pre{\hypertarget{ref-schoneich2020}{}}%
S, S. (2020). Neuroethology of acoustic communication in field
crickets-from signal generation to song recognition in an insect brain.
\emph{Progress in Neurobiology}. Retrieved from
\url{https://www.sciencedirect.com/science/article/pii/S0301008220301374}

\leavevmode\vadjust pre{\hypertarget{ref-Sikes2016}{}}%
Sikes, R.S., Animal Care, the \& American Society of Mammalogists, U.C.
of the. (2016). 2016 guidelines of the american society of mammalogists
for the use of wild mammals in research and education. \emph{Journal of
Mammalogy}, \textbf{97}, 663--688. Retrieved from
\href{https://academic.oup.com/jmammal/article/97/3/663/2459909\%20http://www.mammalsociety.org/committees/animal-care-and-use\#tab3}{https://academic.oup.com/jmammal/article/97/3/663/2459909
http://www.mammalsociety.org/committees/animal-care-and-use\#tab3}

\leavevmode\vadjust pre{\hypertarget{ref-Stowell2022}{}}%
Stowell, D. (2022). Computational bioacoustics with deep learning: A
review and roadmap. \emph{PeerJ}, \textbf{10}, e13152. Retrieved from
\url{https://peerj.com/articles/13152}

\leavevmode\vadjust pre{\hypertarget{ref-sturman2020}{}}%
Sturman, O., Ziegler, L. von, Schläppi, C. \& Akyol, F. (2020). Deep
learning-based behavioral analysis reaches human accuracy and is capable
of outperforming commercial solutions. \emph{Neuropsychopharmacology}.
Retrieved from \url{https://www.nature.com/articles/s41386-020-0776-y}

\leavevmode\vadjust pre{\hypertarget{ref-sugai2019}{}}%
Sugai, L., Silva, T., Jr, J.R. \& Llusia, D. (2019). Terrestrial passive
acoustic monitoring: Review and perspectives. \emph{academic.oup.com}.
Retrieved from
\url{https://academic.oup.com/bioscience/article-abstract/69/1/15/5193506}

\leavevmode\vadjust pre{\hypertarget{ref-Tchernichovski2021}{}}%
Tchernichovski, O., Eisenberg-Edidin, S. \& Jarvis, E.D. (2021).
Balanced imitation sustains song culture in zebra finches. \emph{Nature
Communications 2021 12:1}, \textbf{12}, 1--14. Retrieved from
\url{https://www.nature.com/articles/s41467-021-22852-3}

\leavevmode\vadjust pre{\hypertarget{ref-tiwari2022}{}}%
Tiwari, C. \& Diwakar, S. (2022).
\href{https://doi.org/10.1080/09524622.2022.2064330}{The katydid
country: Bioacoustics and ecology of tettigoniid communities from the
indian subcontinent}. \emph{Bioacoustics}.

\leavevmode\vadjust pre{\hypertarget{ref-zsebok2021}{}}%
Zsebők, S., Schmera, D., Miklós Laczi, \textbar, Nagy, G., Vaskuti, É.,
János Török, \textbar, László, \textbar{} \& Garamszegi, Z. (2021). A
practical approach to measuring the acoustic diversity by community
ecology methods. \emph{Wiley Online Library}, \textbf{12}, 874--884.
Retrieved from
\url{https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/2041-210X.13558}

\end{CSLReferences}

\end{document}