### 2023-05-24 scripts/study_case_zebra_finch_songs.qmd

---
title: Energy-based automatic detection on zebra finch songs
subtitle: ohun
author: <a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a>
date: "`r Sys.Date()`"
toc: true
toc-depth: 2
toc-location: left
number-sections: true
highlight-style: pygments
format:
  html:
    df-print: kable
    code-fold: show
    code-tools: true
    css: qmd.css
---

&nbsp;

<div class="alert alert-info">

### Purpose

- Showcase energy-based automatic detection with the R package ohun using zebra-finch songs

</div>

&nbsp;

<div class="alert alert-warning">

### Data set

The data used in this tutorial (annotations and recordings) can be downloaded from here:

- [https://figshare.com/ndownloader/files/38431340](https://figshare.com/ndownloader/files/38431340)

</div>

&nbsp;

```{r setup, include = FALSE}

knitr::opts_knit$set(root.dir = "..")

knitr::opts_chunk$set(include = TRUE, eval = TRUE, warning = FALSE)

data_path <- raw_data_path <- "data/raw/taeniopygia"

preproc_path <- "data/processed/taeniopygia"

figure_path <- "data/processed/taeniopygia/figures"

source("~/Dropbox/R_package_testing/ohun/R/diagnose_detection.R")
source("~/Dropbox/R_package_testing/ohun/R/label_detection.R")
source("~/Dropbox/R_package_testing/ohun/R/summarize_diagnostic.R")
source("~/Dropbox/R_package_testing/ohun/R/internal_functions.R")
```

### Package installation

Install/load the package from CRAN as follows:

```{r, eval = FALSE}

# From CRAN would be
install.packages("ohun")

#load package
library(ohun)

```

To install the latest developmental version from [github](https://github.com/) you will need the R package [remotes](https://cran.r-project.org/package=remotes):

```{r, eval = FALSE}
remotes::install_github("maRce10/ohun")

#load package
library(ohun)

```

### Load additional packages
```{r libraries, eval = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

x <- c("ohun", "Rraven", "warbleR", "pbapply", "remotes", "DT")

out <- lapply(x, function(y) {

  # check if installed, if not then install 
  if (!y %in% installed.packages()[,"Package"]) 
    install.packages(y)

  # load package
  try(require(y, character.only = T), silent = T)
})

```

```{r, message=FALSE, warning=FALSE}
library(ohun)

library(warbleR)

library(viridis)
```

&nbsp;

### Set directory where the sound files and annotations are found

```{r, eval = FALSE, echo = TRUE}

data_path <- "DIRECTORY_WITH_SOUND_FILES_AND_ANNOTATIONS_HERE"

```

&nbsp;

### Read reference annotations
```{r, eval = TRUE, echo = FALSE}

manual_ref <- read.csv(file.path("data/processed/taeniopygia", "manual_selections_Taeniopygia.csv"))

```

```{r, eval = FALSE, echo = TRUE}

manual_ref <- read.csv(file.path(data_path, "manual_selections_Taeniopygia.csv"))

```

&nbsp;

### Create spectrograms to explore vocalization structure

This code creates a multipanel image with multiple spectrograms, one for each of the individuals/recordings in the complete data set  (note that the output of the function `catalog` are image files saved in the sound file directory):
```{r Create catalogs, eval = FALSE}

# select highest signal to noise ratio calls per individual
manual_ref_snr <- signal_2_noise(X = manual_ref, mar = 0.05, path = data_path)

# select 1 example per sound file
high_snr <- manual_ref_snr[ave(-manual_ref_snr$SNR, manual_ref_snr$sound.files, FUN = rank) <= 1, ]

high_snr$bottom.freq <- -1
high_snr$top.freq <- 12

# create catalogs
catalog(X = high_snr, flim = c(0, 11), nrow = 6, ncol = 3, ovlp = 90, height = 15, width = 20, same.time.scale = TRUE, mar = 0.01, wl = 512, gr = FALSE, spec.mar = 0, lab.mar = 0.001, rm.axes = TRUE, by.row = TRUE, box = TRUE, pal = viridis, parallel = 10, collevels = seq(-140, 0, 5), path = data_path, img.prefix = "same.time", highlight = TRUE, alpha = 0.2)

```

![](img/Catalog_taenopygia.jpeg)

&nbsp;

### Split reference annotations for training and testing detection
```{r trains set, eval = FALSE, echo = FALSE}

# train set must be this one
train_files <- c(
  "BRN7_43435.27985312_12_1_7_46_25.wav",
  "db1_43427.28100313_11_23_7_48_20.wav",
  "Gry35HP_43455.29800260_12_21_8_16_40.wav"
)

```

```{r test and train, eval = TRUE, echo = TRUE}

set.seed(450)    
train_files <- sample(unique(manual_ref$sound.files), 3)

test_files <- setdiff(manual_ref$sound.files, train_files)

train_ref <- manual_ref[manual_ref$sound.files %in% train_files, ]
test_ref <- manual_ref[manual_ref$sound.files %in% test_files, ]
```

&nbsp;

### Optimize detection on training data
```{r  optimize automated detection, eval = FALSE, echo = TRUE}

opt_det <- optimize_energy_detector(
  reference = train_ref, 
  files = train_files, 
  threshold = c(1, 5), 
  hop.size = 11.6, 
  smooth = c(5, 10), 
  hold.time = c(0, 5), 
  min.duration = c(5, 15, 25), 
  max.duration = c(275, 300, 325), 
  bp = c(0.5, 10),
  path = data_path
)

```

```{r  save optimized automated detection, eval = FALSE, echo = FALSE}

write.csv(opt_det, file.path(preproc_path, "detection_optimization.csv"), row.names = FALSE)
```

&nbsp;

### Check subset with highest performance (sorted by F score)
```{r, eval = FALSE}

# subset with highest performance
sub_opt_det <- opt_det[order(opt_det$f.score, decreasing = TRUE), ]

sub_opt_det <- sub_opt_det[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]

#print 
sub_opt_det

```

```{r, eval = TRUE, echo = FALSE}

opt_det <- read.csv(file.path(preproc_path, "detection_optimization.csv"))

# subset with highest performance
sub_opt_det <- opt_det[order(opt_det$f.score, decreasing = TRUE), ]

sub_opt_det <- sub_opt_det[1:10, c("threshold", "smooth", "hold.time", "min.duration", "max.duration", "true.positives", "false.positives",  "false.negatives", "recall", "precision", "f.score")]

# print dynamic table
oa_DT <- datatable(sub_opt_det, editable = list(target = "row"), rownames = FALSE, style = "bootstrap",
    filter = "top", options = list(pageLength = 100, autoWidth = TRUE, dom = "ft"),
    autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(sub_opt_det, is.numeric), 3)
```

&nbsp;

### Select the highest performance parameters based on F score
```{r, select best params, eval = TRUE, echo = TRUE}

best_param <- opt_det[which.max(opt_det$f.score), ]

best_param
```

&nbsp;

### Run detection on test data
```{r, detect on all data, eval = FALSE, echo = TRUE}

det_test <- energy_detector(
  files = test_files, 
  threshold = best_param$threshold, 
  hop.size = 11.6, 
  smooth = best_param$smooth, 
  hold.time = best_param$hold.time, 
  min.duration = best_param$min.duration, 
  max.duration = best_param$max.duration, 
  bp = c(0.5, 10), 
  path = data_path
)
```

```{r , eval = FALSE, echo = FALSE}

write.csv(det_test, file.path(preproc_path, "detection_test_data.csv"), row.names = FALSE)

```

&nbsp;

### Evaluate detection performance

```{r , eval = TRUE, echo = FALSE}

det_test <- read.csv(file.path(preproc_path, "detection_test_data.csv"))

```

```{r, diagnose, eval = TRUE, echo = TRUE}

diagnose_detection(reference = test_ref, detection = det_test, by.sound.file = FALSE)

```

&nbsp;

<div class="alert alert-info">

### Takeaways

- Good performance on detecting zebra finch songs: F score was 0.95 for the training data set and 0.96 for the testing data

</div>

&nbsp;

---

<font size="4">Session information</font>

```{r session info, echo=F}

sessionInfo()

```